# Basic UI 

```{r}
library(tidyverse)
library(magrittr)
library(shiny)
```

## Free Text 

```{r}
ui <- fluidPage(
    # collect small amounts of text
    textInput("name", "What's your name?"),
    # enter passwords
    passwordInput("password", "What's your password?"),
    # paragraphs of text
    textAreaInput("story", "Tell me about yourself", rows = 3)    
)
```

If we want to ensure that text has certain properties we can use validate(). 

## Numeric Inputs 

```{r}
ui <- fluidPage(
    # collect numbers in a constrained textbox
    numericInput("num", "Number One", value = 0, min = 0, max = 100),
    # number slider
    sliderInput("num2", "Number Two", value = 50, min = 0, max = 100),
    # double ended number slider (range)
    sliderInput("num3", "Range", value = c(10, 20), min = 0, max = 100)
)
```

## Dates 

```{r}
ui <- fluidPage(
    dateInput("dob", "When were you born?"),
    dateRangeInput("holiday", "When do you want to go on vacation next?")
)
```

Arguments like datesdisabled and daysofweekdisabled allow us to restrict the set of valid inputs.

## Limited Choices 

```{r}
animals <- c("dog", "cat", "mouse", "bird", "other")

ui <- fluidPage(
    # input selector
    selectInput("state", "What's your favorite state?", state.name),
    # radio button choices
    radioButtons("animal", "What's your favorite animal?", animals)
)
```


Radio buttons can also show options other than plain text 

```{r}
ui <- fluidPage(
    radioButtons("rb", "Choose One:",
                 choiceNames = list(
                     icon("angry"),
                     icon("smile"),
                     icon("sad-tear")),
                 choiceValues = list("angry", "happy", "sad")))
```

With selectInputs we can also allow for multiple choices 

```{r}
ui <- fluidPage(selectInput("state", "What's your favorite state?",
                            state.name, multiple = TRUE))
```

The multiple alternative for radio buttons is the check box

```{r}
ui <- fluidPage(checkboxGroupInput("animal", "What animals do you like?", animals))
```

If we want a single checkbox for a single yes/no question, use checkboxInput()

```{r}
ui <- fluidPage(
    checkboxInput("cleanup", "Clean Up?", value = TRUE),
    checkboxInput("shutdown", "Shutdown?"))
```

## File Uploads 

```{r}
ui <- fluidPage(fileInput("upload", NULL))
```

This requires special handling on the server side and is covered later. 

## Action Buttons 

```{r}
ui <- fluidPage(
    actionButton("click", "Click me!"),
    actionButton("drink", "Drink me!", icon = icon("cocktail")))
```

## Exercises 

1. When space is at a premium, its useful to label text boxes using a placeholder that appears inside the text entry area. 

```{r}
textInput("entry", placeholder = "Your Name")
```

2. 

```{r}
sliderInput("name", "When should we deliver?", timeFormat = "%F")
```

3. 

We create subheadings by breaking the choices into groups

```{r}
selectInput("state", "Choose a state:",
            list(`East Coast` = list("NY", "NJ", "CT"),
                 `West Coast` = list("WA", "OR", "CA"),
                 `Midwest` = list("MN", "WI", "IA")))
```

4. 

```{r}
sliderInput("name", "Label", min = 0, max = 100, value = 50, step = 5,
            animate = TRUE)
```

5. The step parameter changes how the values move upwards and downwards when the side arrow keys are clicked. 

# Outputs

Each output function on the front end is coupled with a render function in the back end. There are 3 main types of output, corresponding to the 3 things we usually include in a report:

1. text 
2. tables 
3. plots 

## Text 

Output regular text with textOutput() and fixed code and console output with verbatimTextOutput()

```{r}
ui <- fluidPage(
    textOutput("text"),
    verbatimTextOutput("code")
)

server <- function(input, output, session) {
    output$text <- renderText({
        "Hello, friend!"
    })

    output$code <- renderPrint({
        summary(1:10)
    })
}

shinyApp(ui, server)
```

The parenthesis are only needed when we wish to run multiple lines of code. If we have a one liner (like above), we could omit them. 

There are two render functions that can be used with either of the text output functions:

1. renderText() displays text returned by the code 
2. renderPrint() displays text printed by the code 

```{r}
print_and_return <- function() {
    print("a")
    print("b")
    "c"
}

x <- print_and_return()
x
```

## Tables 

