# Case Study 

```{r}
library(shiny)
library(tidyverse)
library(vroom)
```

## Data 

```{r}
# get data
nice <- neiss::injuries %>%
    filter(lubridate::year(trmt_date) == 2017)

prods <- neiss::products

popn <- neiss::population

# write to csv
list(nice, prods, popn) %>%
    set_names(c("nice", "prods", "popn")) %>%
    imap(., ~ write_csv(x = .x, path = paste0("neiss_", .y, ".csv")))

# load with vroom
nice <- vroom("neiss_nice.csv", delim = ",")
prods <- vroom("neiss_prods.csv", delim = ",")
popn <- vroom("neiss_popn.csv", delim = ",")
```

- trmt_date is the date the person was seen in the hospital 
- age, sex, race give demographic information 
- body_part is the location of the injury on the body
- location is where the accident occurred 
- diag gives the basic diagnosis of the injury 
- prod_code is the primary product associated with the injury 
- weight is the statistical weight given the estimated number of people who would suffer from this injury if the dataset was scaled to the entire population of the us 
- narrative is a brief story about how the accident occurred 

- products lets us look up the product name from the product code 
- population tells us the total us population in that year for each combination of age and sex 

## Exploration

What is the product with the most injuries?

```{r}
nice %>%
    group_by(prod1) %>%
    tally(sort = TRUE) %>%
    mutate(propn = round(n / sum(n), 2))

(nice %>%
    filter(prod1 == 1842) %>%
    left_join(prods, by = c("prod1" = "code")) %>%
    select(title, everything()) -> selected)
```

stairs or steps, with 8% by proportion. Following that up, also with 8%, is floors or flooring materials. 


What are some basic summaries of diagnosis, body part and location weighted by the weight variable?

```{r}
weighted_count <- function(data, variable) {
    require(rlang)

    data %>%
        count(eval(parse_expr(variable)), wt = weight, sort = TRUE)
}


c("diag", "body_part", "location") %>%
    map(., ~ weighted_count(selected, .x)) %>%
    set_names(c("diag", "body_part", "location"))
```


steps injuries most commonly are associated with strain or sprains of the ankle and often happen at home. 

```{r}
selected %>%
    group_by(age, sex) %>%
    tally(sort = TRUE, wt = weight) %>%
    ggplot(aes(age, n, color = sex)) +
    geom_line() +
    labs(y = "Estimated Number of Injuries")
```

We see a big spike when children are learning to walk, a drop off and then a reduction after 50. Women tend to have more steps related injuries. This might be due to high heeled shoes

One problem with interpreting this pattern is that we know that there are fewer older people than younger people. We can control for this by comparing the number of people injured with the total population and calculating an injury rate. 

```{r}

selected %>%
    group_by(age, sex) %>%
    tally(sort = TRUE) %>%
    mutate(sex = tolower(sex)) %>%
    left_join(popn %>%
              filter(year == 2017) %>%
              select("population" = n, everything()),
              by = c("age", "sex")) %>%
    mutate(rate = (n / population) * 1e5) -> summary

nice %>%
    filter(prod1 == 1207) %>%
    mutate(sex = tolower(sex)) %>%
    group_by(age, sex) %>%
    tally(sort = TRUE) %>%
    left_join(popn %>%
              filter(year == 2017) %>%
              select("population" = n, everything()),
              by = c("age", "sex")) %>%
    mutate(rate = (n / population) * 1e5)

summary %>%
    ggplot(aes(age, rate, color = sex)) +
    geom_line() +
    labs(y = "Estimated Number of Injuries per 10,000 People")
```

Now we can look at some narratives

```{r}
selected %>%
    sample_n(10) %>%
    pull(narrative)
```

Having done this analysis for one product, we would like to easily do it for other products. This is a good use case for a shiny app

# Prototype 

When building a complex app, it is recommended to start as simply as possible to confirm the basic mechanics work before doing something more complicated.

Here is a beginner design with one row for inputs, one row for all three tables, and one row for the plot

```{r}

ui <- fluidPage(
    fluidRow(
        column(6,
               selectInput("code", "Product",
                           setNames(prods$code, prods$title)))),
    fluidRow(
        column(4, tableOutput("diag")),
        column(4, tableOutput("body_part")),
        column(4, tableOutput("location"))),
    fluidRow(
        column(12, plotOutput("age_sex"))))

```

The server function is straight forward. First we convert the selected and summary variables to reactive expressions. 

```{r}
server <- function(input, output, session) {
    
    selected <- reactive(nice %>% filter(prod1 == input$code))

    output$diag <- renderTable(
        selected() %>% count(diag, wt = weight, sort = TRUE)
    )
    output$body_part <- renderTable(
        selected() %>% count(body_part, wt = weight, sort = TRUE)
    )
    output$location <- renderTable(
        selected() %>% count(location, wt = weight, sort = TRUE))

    summary <- reactive({
        selected() %>%
            mutate(sex = tolower(sex)) %>%
            group_by(age, sex) %>%
            tally(sort = TRUE) %>%
            left_join(popn %>%
                      filter(year == 2017) %>%
                      select("population" = n, everything()),
                      by = c("age", "sex")) %>%
            mutate(rate = (n / population) * 1e5)
    })

    output$age_sex <- renderPlot({
        summary() %>%
            ggplot(aes(age, n, color = sex)) +
            geom_line() +
            labs(y = "Estimated Number of Injuries") +
            theme_grey(15)})}

shinyApp(ui, server)
```

## Polish Tables 

The first problem with the app is that it shows a lot of information in the tables, where we probably just want highlights.

We can turn the variables into factors, order by frequency, and lump together the leftovers after the top 5.

```{r}
count_top <- function(df, var, n = 5) {
    df %>%
        mutate({{var}} := fct_lump(fct_infreq({{var}}), n = n)) %>%
        group_by({{var}}) %>%
        summarize(n = as.integer(sum(weight)))
}
```

Now we can update the server function 

```{r}
server <- function(input, output, session) {
       selected <- reactive(nice %>% filter(prod1 == input$code))

       output$diag <- renderTable(count_top(selected(), diag), width = "100%")
       output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
       output$location <- renderTable(count_top(selected(), location), width = "100%")

       summary <- reactive({
           selected() %>%
               mutate(sex = tolower(sex)) %>%
               group_by(age, sex) %>%
               tally(sort = TRUE) %>%
               left_join(popn %>%
                         filter(year == 2017) %>%
                         select("population" = n, everything()),
                         by = c("age", "sex")) %>%
               mutate(rate = (n / population) * 1e4)})

       output$age_sex <- renderPlot({
           summary() %>%
               ggplot(aes(age, n, color = sex)) +
               geom_line() +
               labs(y = "Estimated Number of Injuries") +
               theme_grey(15)})
}

shinyApp(ui, server)
```

## Rate vs Count

We wish to give users the choice between visualizing the number of injuries or the population standardized rate.

First we add a control to the UI:

```{r}

ui <- fluidPage(
    fluidRow(
        column(6,
               selectInput("code", "Product",
                           setNames(prods$code, prods$title),
                           width = "100%")),
        column(2, selectInput("y", "Y axis", c("rate", "count")))),
    fluidRow(
        column(4, tableOutput("diag")),
        column(4, tableOutput("body_part")),
        column(4, tableOutput("location"))),
    fluidRow(
        column(12, plotOutput("age_sex"))))

```

Then we condition on that input when generating the plot 

```{r}

server <- function(input, output, session) {
       selected <- reactive(nice %>% filter(prod1 == input$code))

       output$diag <- renderTable(count_top(selected(), diag), width = "100%")
       output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
       output$location <- renderTable(count_top(selected(), location), width = "100%")

       summary <- reactive({
           selected() %>%
               mutate(sex = tolower(sex)) %>%
               group_by(age, sex) %>%
               tally(sort = TRUE) %>%
               left_join(popn %>%
                         filter(year == 2017) %>%
                         select("population" = n, everything()),
                         by = c("age", "sex")) %>%
               mutate(rate = (n / population) * 1e5)})

       output$age_sex <- renderPlot({
           if (input$y == "count") {
               summary() %>%
                   ggplot(aes(age, n, color = sex)) +
                   geom_line() +
                   labs(y = "Estimated Number of Injuries") +
                   theme_grey(15)
           } else {
               summary() %>%
                   ggplot(aes(age, rate, color = sex)) +
                   geom_line() +
                   labs(y = "Estimated Number of Injuries per 10,000 People") +
                   theme_grey(15)}})}

shinyApp(ui, server)
```


## Narrative

We wish to add some of the narratives because they give perspective to the plots and tables.

First we want to add a new row to the bottom of the UI. Then we can add an action button to trigger a new story, and put the narrative in a textOutput

```{r}
ui <- fluidPage(
    fluidRow(
        column(6,
               selectInput("code", "Product",
                           setNames(prods$code, prods$title),
                           width = "100%")),
        column(2, selectInput("y", "Y axis", c("rate", "count")))),
    fluidRow(
        column(4, tableOutput("diag")),
        column(4, tableOutput("body_part")),
        column(4, tableOutput("location"))),
    fluidRow(
        column(12, plotlyOutput("age_sex"))),
    fluidRow(
        column(2, actionButton("story", "Tell Me a Story")),
        column(10, textOutput("narrative"))))
```

The result of an action button is an integer that increments each time it is clicked. Here we use it to trigger a re-execution of the random selection.

```{r}
server <- function(input, output, session) {
       selected <- reactive(nice %>% filter(prod1 == input$code))

       output$diag <- renderTable(count_top(selected(), diag), width = "100%")
       output$body_part <- renderTable(count_top(selected(), body_part), width = "100%")
       output$location <- renderTable(count_top(selected(), location), width = "100%")

       summary <- reactive({
           selected() %>%
               mutate(sex = tolower(sex)) %>%
               group_by(age, sex) %>%
               tally(sort = TRUE) %>%
               left_join(popn %>%
                         filter(year == 2017) %>%
                         select("population" = n, everything()),
                         by = c("age", "sex")) %>%
               mutate(rate = (n / population) * 1e5)})

       output$age_sex <- renderPlotly({
           if (input$y == "count") {
               summary() %>%
                   ggplot(aes(age, n, color = sex)) +
                   geom_line() +
                   labs(y = "Estimated Number of Injuries") +
                   theme_grey(15) -> p1 
                   p1 %>% plotly::ggplotly()
           } else {
               summary() %>%
                   ggplot(aes(age, rate, color = sex)) +
                   geom_line() +
                   labs(y = "Estimated Number of Injuries per 10,000 People") +
                   theme_grey(15) -> p1
                   p1 %>% plotly::ggplotly()}})

       output$narrative <- renderText({
           input$story
           selected() %>%
               pull(narrative) %>%
               sample(1)})
}

shinyApp(ui, server)
```
