# Reducing Duplication

```{r}
library(tidyverse)
library(shiny)

sp <- function(){
    shinyApp(ui, server)
}
```

Every piece of knowledge must have a single, unambiguous authoritive representation within a system.

Advantages of functions and modules:

- clearly isolated behavior through specified inputs and outputs means that it is easier to understand how parts of the app fit together, so we don't need to worry about spooky action at a distance where changing one part of the app changes the way an apparently unrelated part works
- reducing duplication makes it easier to respond to changing needs because instead of having to track down and change every piece of duplicated code, we can just change it in one place 
- we can spread our app across multiple files so that it can be more easily digested in chunks. 

# Using Functions

UI Code or self contained reactives are a good use case for functions. A function will only take you so far though because we typically need to have some connection between the front end and the back end, and we need some way to coordinate the two. Shiny uses identifiers, so we need some way to share them. This gives rise to shiny **modules**.

## Helper Functions

If, given specific values, our app requires a complex calculation, we should pull it out into a separate function. We want to go from this:

```{r}
server <- function() {
    data <- reactive({
        # some complex data calculation involving inputs x, y, z
})}
```

to this:

```{r}
my_helper <- function(x, y, z) {
    # ...
}

server <- function() {
    data <- reactive({
        my_helper(input$x, input$y, input$z)})}
```

When extracting out such helpers, avoid putting any reactive component inside the function. Instead pass them in through the arguments. 

Advantages of functions: 

- it allows us to move it to a separate file 
- it makes it clear from the outside exactly what the function takes for inputs. When looking at a reactive expression or output, there is no easy way to tell exactly what values it depends on without carefully reading the code block
- it enforces independence - if we try and refer to an input that was not passed to the function, we get an error. This becauses increasingly important as our apps get bigger because it ensures that the pieces of the app and independent and can be analyzed in isolation

## UI Functions

```{r}
sliderInput01 <- function(id, label = id) {
    sliderInput(id, label,
                min = 0, max = 1, value = 0.5, step = 0.1)
}

ui <- fluidRow(
    sliderInput01("alpha"),
    sliderInput01("beta"),
    sliderInput01("gamma"),
    sliderInput01("delta")
)
```

We could reduce it more with functional programming.

fluidRow(!!!list(a, b)) is equivalent to fluidRow(a, b). This technique is sometimes called splatting because you're splatting the elements of a list into the arguments of a function. 

```{r}
vars <- c("alpha", "beta", "gamma", "delta")
sliders <- map(vars, sliderInput01)
ui <- fluidRow(!!!sliders)
```

## Reactives 

We wish to keep as much inside the server function as possible.

## Case Study 

There was a post on the rstudio community site where the user wanted to add multiple select inputs in the sidebar and in the main body and have the reactive graph change when any of the inputs are selected or changed.

The UI looked something like this

```{r}
fluidRow(
    box(
        width = 4,
        solidHeader = TRUE,
        selectInput("traffickingType",
                    label = "Choose a trafficking type: ",
                    choices = sort(unique(ngo$Trafficking.Type)),
                    multiple = TRUE)),
    box(
        width = 4,
        solidHeader = TRUE,
        selectInput("traffickingSubType",
                    label = "Choose a trafficking sub type: ",
                    choices = sort(unique(ngo$Trafficking.Sub.Type)),
                    multiple = TRUE)),
        box(
        width = 4,
        solidHeader = TRUE,
        selectInput("gender",
                    label = "Choose a gender: ",
                    choices = sort(unique(ngo$Victim.Gender)),
                    multiple = TRUE)),
)
```

There is a lot of duplication here. 

- The call to box repeats with some of the same arguments over and over. It appears the intent of the code is to make a header, so we can call the function headerBox. 
- The calls to selectInput return multiple = TRUE and all use the same strategy for determining the choices: pulling unique values from a data frame column. This function is tied to a specific dataset, so we can call it ngoSelectInput(). 

```{r}
ngo_select_input <- function(var, label, multiple = TRUE) {
    choices <- sort(unique(ngo[[var]]))
    label <- paste0("Choose a ", label, ": ")
    selectInput(var, label, choices = choices, multiple = multiple)
}

box_header <- function(...) {
    box(width = 4, solidHeader =  TRUE, ...)
}

fluidRow(box_header(ngo_select_input("Trafficking.Type", "trafficking type")),
         box_header(ngo_select_input("Trafficking.Sub.Type", "trafficking sub type")),
         box_header(ngo_select_input("Victim.Gender", "gender")))
```

We made an assumption that would also require changes on the server side: when filtering based on a variable, the input name should be the same as the variable name. This is generally a good idea, as it makes the code easier to reason about 

We could also generalize this to handle multiple datasets

```{r}
dfSelectInput <- function(df, var, label, multiple = TRUE) {
    choices <- sort(unique(df[[var]]))
    label <- paste0("Choose a ", label, ": ")
    selectInput(var, label, choices = choices, multiple = multiple)
}
```

This is a good idea if we saw that pattern repeated in multiple places -- but we probably also need to introduce some additional components for the ID, otherwise we risk generating controls with the same id. This is the problem of namespacing -- we want to have a hierarchy of names. This is one of the big problems that modules solve.

If we had a lot of variables, we could use functional programming to generate them. The key idea is to capture all the data needed to generate the columns in a single data frame. Dataframes are useful because they gemeralize to any number of arguments.

Here is an example: 

```{r}
vars <- tibble::tribble(~ var, ~label,
                        "t_type", "trafficking type",
                        "t_sub_type", "trafficking sub type",
                        "victim_gender", "gender")

# use pmap to turn each row in the df to a call to ngoSelectInput
vars %>%
    pmap(ngo_select_input) %>% # create a select input for each row 
    map(box_header) %>% # wrap each in a box header
    fluidRow(!!!.) # splat into a fluidRow 
```

we could be even more advanced and generate the call to dplyr::filter

```{r}
# create filter calls
(select <- map(vars$var, function(v) rlang::expr(.data[[!!v]] == input[[!!v]])))

# splat
filter(ngo, !!!select)
```

# Using Modules

functions are great but they are only useful for extracting pure ui code or pure computation used inside reactives. They don't help if we wish to build more complicated components that link ui and server.

A shiny module is a pair of functions corresponding to the front end UI and the backend server function. 

They essentially allow us to create an app within an app. They force isolation of behavior so there is no stepping on toes.

## Without Modules

As an example of why we need modules -- consider an app that has the user input a date as a string and validate it 

```{r}
library(lubridate)

ui <- fluidPage(
    textInput("date", "When were you born? (yyyy-mm-dd)"),
    textOutput("error"),
    textOutput("age"))

server <- function(input, output, session) {
    birthday <- reactive({
        req(input$date)
        ymd(input$date, quiet = TRUE)})

    age <- reactive({
        req(birthday())
        (birthday() %--% today() %/% years(1))})
    
    output$error <- renderText({
        if (is.na(birthday())) "Please enter valid date in yyyy-mm-dd form"})

    output$age <- renderText({paste0("You are ", age(), "years old")})
}

shinyApp(ui, server)
```

It is plausible that as the app gets bigger we might want to use the date control in multiple places -- so lets extract it into functions

```{r}
ymdInputUI <- function(label) {
    label <- paste0(label, " (yyyy-mm-dd)")

    fluidRow(
        textInput("date", label),
        textOutput("error"))}

ymdInputServer <- function(input, output, session) {
    date <- reactive({
        req(input$date)
        ymd(input$date, quiet = TRUE)})

    output$error <- renderText({
        if (is.na(date())) "Please enter valid date in yyyy-mm-dd form"})

    date
}

ui <- fluidPage(
    ymdInputUI("When were you born?"),
    textOutput("age"))

server <- function(input, output, session) {
    birthday <- ymdInputServer(input, output, session)

    age <- reactive({
        req(birthday())
        (birthday() %--% today() %/% years(1))})

    output$age <- renderText({
        paste0("You are ", age(), " years old")})}
```


There are two problems with this app: 

- it always assumes that the control is called date. This means we can't have two controls in the same app.
- the ui has an output with id error that you can't see from reading just the UI code. This makess it easy to break the app 

```{r}
ui <- fluidPage(
    ymdInputUI("When were you born?"),
    textOutput("error")
)
```

These problems arise because we've used functions to isolate local variables. The code is simpler to understand because we can't access the internals of the functions -- but there is another way that shiny code can interface: through the names of input and output controls.

This is the problem modules are made to solve. 

## Making a Module 

To convert the code above into a module, we need to make two changes: 

1. add an id argument to our UI component
2. use it with a special NS() function. This creates a namespace, or space of names, that is unique to the module 

```{r}
ymdInputUI <- function(id, label) {
    ns <- NS(id)
    label <- paste0(label, " (yyyy-mm-dd)")

    fluidRow(
        textInput(ns("date"), label),
        textOutput(ns("error")))}
```

The key idea is that the argument to NS() is supplied by the person using the component, and the args to the function it produces is supplied by the person who wrote the component. This ensures that the final name combines properties needed by both the app and module author. 

We now need to specify an id when creating the UI. This is important because it puts this id in the same place as all the others, so it is easy to spot if we've used the same input id in multiple places.

```{r}
ui <- fluidPage(
    ymdInputUI("birthday", "When were you born?"),
    textOutput("age")
)
```

Now we make a similar change to the server side of the module. Here, instead of NS, we use callModule which automatically tweaks the input and output so it looks for date inside the id namespace. Its doing something like input[[paste(id, "-", birthday)]]. 

```{r}
ymdInput <- function(id) {
    callModule(id = id, function(input, output, session) {
        date <- reactive({
            req(input$date)
            ymd(input$date, quiet = TRUE)})

        output$error <- renderText({
            if (is.na(date())) "Please enter valid date in yyyy-mm-dd form"})

        date
    })
}
```

Now the arguments to ymdInput have changed -- we pass in the id and shiny takes care of automatically plumbing up the input, output and session in the appropriate namespaced way. 

```{r}
server <- function(input, output, session) {
    birthday <- ymdInput("birthday")

    age <- reactive({
        req(birthday())
        (birthday() %--% today()) %/% years(1)})

    output$age <- renderText({
        paste0("You are ", age(), " years old")})}
```

## Limited Selection + Other

Consider the following app which provides a way to select gender that is sensitive to the many possible ways that people can express their gender 

```{r}
ui <- fluidPage(
    radioButtons("gender", "Gender:",
                 choiceValues = list("male", "female", "self-described", "na"),
                 choiceNames = list("Male", "Female",
                                textInput("gender_self", NULL,
                                          placeholder = "Self-Described"),
                                "Prefer not to say"),
                 selected = "na"),
    textOutput("txt"))

server <- function(input, output, session) {
    observeEvent(input$gender_self, {
        req(input$gender_self)
        updateRadioButtons(session, "gender",
                           selected = "self-described")})

    gender <- reactive({
        if (input$gender == "self-described") {
            input$gender_self
        } else {
            input$gender
        }})

    output$txt <- renderText({
        paste("You chose", gender())})}

sp()
```

Now, converting to a module and generalizing a little 

```{r}
radioButtonsWithOther <- function(id, label, choices,
                                  selected = NULL, placeholder = NULL) {
    ns <- NS(id)

    radioButtons(ns("primary"), "Gender:",
                 choiceValues = c(names(choices), "other"),
                 choiceNames = c(
                     unname(choices),
                     list(textInput(ns("other"),
                                    NULL,
                                    placeholder = NULL))),
                     selected = selected)
}

radioButtonsWithOtherServer <- function(input, output, session) {
    observeEvent(input$primary, {
        req(input$other)
        updateRadioButtons(session, "primary", selected = "other")})

    reactive({
        if (input$primary == "other") {
            input$other
        } else {
            input$primary
        }})}
```

```{r}
ui <- fluidPage(
    radioButtonsWithOther("gender",
                          choices = list(
                              male = "Male",
                              female = "Female",
                              na = "Prefer not to say"),
                          placeholder = "Self-Described",
                          selected = "na"),
    textOutput("txt"))

server <- function(input, output, session) {
    gender <- callModule(radioButtonsWithOtherServer, "gender")

    output$txt <- renderText(paste("You chose", gender()))
}

sp()
```

## Hierarchical Select Boxes

```{r}
country_df <- countrycode::codelist %>%
    as_tibble() %>%
    select(iso3c, continent, country = cow.name) %>%
    filter(!is.na(continent), !is.na(country))

continents <- country_df$continent %>% unique() %>% sort()

ui <- fluidPage(
    selectInput("continent", "Continent", choices = continents),
    selectInput("country", "Country", choices = NULL))

server <- function(input, output, session) {
    countries <- reactive({
        country_df[country_df$continent == input$continent, ,drop = FALSE]})

    observeEvent(input$continent, {
        updateSelectInput(session, "country", choice = countries()$country)})}

sp()
```

## Modal

```{r}
modalContinue <- function(id) {
    ns <- NS(id)

    showModal(
        modalDialog(
            "This is really important!",
            "Are you sure you want to continue?",
            footer = list(
                actionButton(ns("yes"), "yes"),
                actionButton(ns("no"), "no")),
            easyClose = FALSE))}

modalContinueServer <- function(id) {
    module <- function(input, output, session) {
        val <- reactiveVal(NULL)

        observeEvent(input$yes, {
            removeModal()
            val(TRUE)})

        observeEvent(input$no, {
            removeModal()
            val(FALSE)})

        reactive(val())}
    callModule(module, id)}
```

```{r}
ui <- fluidPage(
    actionButton("go", "Go"),
    textOutput("result"))

server <- function(input, output, session) {
    observeEvent(input$go, {
        modalContinue("ok")})

    result <- modalContinueServer("ok")

    output$result <- renderText(result())
}

sp()
```

# Exercises

1. The following app plots user selected variables from the msleep dataset for three different types of mammals (carnivores, omnivores, herbavores), with one tab for each type of mammal. Remove the redundancy in the selectInput definitions with the use of functions

```{r}

selectInputMammal <- function(id, label, selected = "sleep_rem") {
    ns <- NS(id)
    
    choices <- c("sleep_total", "sleep_rem", "sleep_cycle",
                 "awake", "brainwt", "bodywt")
    
    selectInput(inputId = ns("mammal"),
                label = label,
                choices = choices,
                selected = selected)
}

ui <- fluidPage(
    selectInputMammal("x", "X-axis"),
    selectInputMammal("y", "Y-axis", selected = "sleep_total"),
    tabsetPanel(id = "vore",
                tabPanel("Carnivore",
                         plotOutput("plot_carni")),
                tabPanel("Omnivore",
                         plotOutput("plot_omni")),
                tabPanel("Herbivore",
                         plotOutput("plot_herbi"))))
```

2. Further remove redundancy by modularizing how subsets and plots are created 

```{r}
# UI


server <- function(input, output, session) {
    # make subsets
    carni <- mammalSelectInput("animal", "carni")
    output$plot_carni <- make_plot_output(carni(), input$x, input$y)
    
    omni <- mammalSelectInput("animal", "omni")
    output$plot_omni <- make_plot_output(omni(), input$x, input$y)
    
    herbi <- mammalSelectInput("animal", "herbi")
    output$plot_herbi <- make_plot_output(herbi(), input$x, input$y)
}

sp()
```

### Extra Examples of Modules

A small but complete application demonstrating a simple "counter" module

```{r}
# module UI
counterButton <- function(id, label = "Counter") {
    # initialize namespace for ui. This returns input[[paste(id, "-", "counter")]]
    # This encapsulates the modules UI
    ns <- NS(id)
    
    tagList(
        actionButton(ns("button"), label = label),
        verbatimTextOutput(ns("out")))}

selectInputMammal <- function(id, label, selected = "sleep_rem") {
    ns <- NS(id)
    
    choices <- c("sleep_total", "sleep_rem", "sleep_cycle",
                 "awake", "brainwt", "bodywt")
    
    selectInput(inputId = ns("mammal"),
                label = label,
                choices = choices,
                selected = selected)
}

# module server side logic
counter <- function(input, output, session) {
    # initialize counter
    count <- reactiveVal(0)

    observeEvent(input$button, {
        count(count() + 1)
    })

    
    output$out <- renderText(count())

    # return a reactive value
    count
}

subset_and_plot_mammal <- function(input, output, session) {
    
}

ui <- fluidPage(counterButton("counter1", "Counter #1"))

server <- function(input, output, session) {
    # callmodule initializes the module 
    callModule(counter, "counter1")
}

sp()
```

```{r}
library(tidyverse)

ui <- fluidPage(
  selectInput(inputId = "x",
              label = "X-axis:",
              choices = c("sleep_total", "sleep_rem", "sleep_cycle", 
                          "awake", "brainwt", "bodywt"),
              selected = "sleep_rem"),
  selectInput(inputId = "y",
              label = "Y-axis:",
              choices = c("sleep_total", "sleep_rem", "sleep_cycle", 
                          "awake", "brainwt", "bodywt"),
              selected = "sleep_total"),
  tabsetPanel(id = "vore",
              tabPanel("Carnivore",
                       plotOutput("plot_carni")),
              tabPanel("Omnivore",
                       plotOutput("plot_omni")),
              tabPanel("Herbivore",
                       plotOutput("plot_herbi")))
)

server <- function(input, output, session) {

  # make subsets
  carni <- reactive( filter(msleep, vore == "carni") )
  omni  <- reactive( filter(msleep, vore == "omni")  )
  herbi <- reactive( filter(msleep, vore == "herbi") )

  # make plots
  output$plot_carni <- renderPlot({
    ggplot(data = carni(), aes_string(x = input$x, y = input$y)) +
      geom_point()
  })
  output$plot_omni <- renderPlot({
    ggplot(data = omni(), aes_string(x = input$x, y = input$y)) +
      geom_point()
  })
  output$plot_herbi <- renderPlot({
    ggplot(data = herbi(), aes_string(x = input$x, y = input$y)) +
      geom_point()
  })

}

shinyApp(ui = ui, server = server)
```

