# Reducing Duplication

```{r}
library(tidyverse)
library(shiny)

sp <- function(){
    shinyApp(ui, server)
}
```

Every piece of knowledge must have a single, unambiguous authoritive representation within a system.

Advantages of functions and modules:

- clearly isolated behavior through specified inputs and outputs means that it is easier to understand how parts of the app fit together, so we don't need to worry about spooky action at a distance where changing one part of the app changes the way an apparently unrelated part works
- reducing duplication makes it easier to respond to changing needs because instead of having to track down and change every piece of duplicated code, we can just change it in one place 
- we can spread our app across multiple files so that it can be more easily digested in chunks. 

# Using Functions

UI Code or self contained reactives are a good use case for functions. A function will only take you so far though because we typically need to have some connection between the front end and the back end, and we need some way to coordinate the two. Shiny uses identifiers, so we need some way to share them. This gives rise to shiny **modules**.

## Helper Functions

If, given specific values, our app requires a complex calculation, we should pull it out into a separate function. We want to go from this:

```{r}
server <- function() {
    data <- reactive({
        # some complex data calculation involving inputs x, y, z
})}
```

to this:

```{r}
my_helper <- function(x, y, z) {
    # ...
}

server <- function() {
    data <- reactive({
        my_helper(input$x, input$y, input$z)})}
```

When extracting out such helpers, avoid putting any reactive component inside the function. Instead pass them in through the arguments. 

Advantages of functions: 

- it allows us to move it to a separate file 
- it makes it clear from the outside exactly what the function takes for inputs. When looking at a reactive expression or output, there is no easy way to tell exactly what values it depends on without carefully reading the code block
- it enforces independence - if we try and refer to an input that was not passed to the function, we get an error. This becauses increasingly important as our apps get bigger because it ensures that the pieces of the app and independent and can be analyzed in isolation

## UI Functions

```{r}
sliderInput01 <- function(id, label = id) {
    sliderInput(id, label,
                min = 0, max = 1, value = 0.5, step = 0.1)
}

ui <- fluidRow(
    sliderInput01("alpha"),
    sliderInput01("beta"),
    sliderInput01("gamma"),
    sliderInput01("delta")
)
```

We could reduce it more with functional programming.

fluidRow(!!!list(a, b)) is equivalent to fluidRow(a, b). This technique is sometimes called splatting because you're splatting the elements of a list into the arguments of a function. 

```{r}
vars <- c("alpha", "beta", "gamma", "delta")
sliders <- map(vars, sliderInput01)
ui <- fluidRow(!!!sliders)
```

## Reactives 

We wish to keep as much inside the server function as possible.

## Case Study 

There was a post on the rstudio community site where the user wanted to add multiple select inputs in the sidebar and in the main body and have the reactive graph change when any of the inputs are selected or changed.

The UI looked something like this

```{r}
fluidRow(
    box(
        width = 4,
        solidHeader = TRUE,
        selectInput("traffickingType",
                    label = "Choose a trafficking type: ",
                    choices = sort(unique(ngo$Trafficking.Type)),
                    multiple = TRUE)),
    box(
        width = 4,
        solidHeader = TRUE,
        selectInput("traffickingSubType",
                    label = "Choose a trafficking sub type: ",
                    choices = sort(unique(ngo$Trafficking.Sub.Type)),
                    multiple = TRUE)),
        box(
        width = 4,
        solidHeader = TRUE,
        selectInput("gender",
                    label = "Choose a gender: ",
                    choices = sort(unique(ngo$Victim.Gender)),
                    multiple = TRUE)),
)
```

There is a lot of duplication here. 

- The call to box repeats with some of the same arguments over and over. It appears the intent of the code is to make a header, so we can call the function headerBox. 
- The calls to selectInput return multiple = TRUE and all use the same strategy for determining the choices: pulling unique values from a data frame column. This function is tied to a specific dataset, so we can call it ngoSelectInput(). 

```{r}
ngo_select_input <- function(var, label, multiple = TRUE) {
    choices <- sort(unique(ngo[[var]]))
    label <- paste0("Choose a ", label, ": ")
    selectInput(var, label, choices = choices, multiple = multiple)
}

box_header <- function(...) {
    box(width = 4, solidHeader =  TRUE, ...)
}

fluidRow(box_header(ngo_select_input("Trafficking.Type", "trafficking type")),
         box_header(ngo_select_input("Trafficking.Sub.Type", "trafficking sub type")),
         box_header(ngo_select_input("Victim.Gender", "gender")))
```

We made an assumption that would also require changes on the server side: when filtering based on a variable, the input name should be the same as the variable name. This is generally a good idea, as it makes the code easier to reason about 

We could also generalize this to handle multiple datasets

```{r}
dfSelectInput <- function(df, var, label, multiple = TRUE) {
    choices <- sort(unique(df[[var]]))
    label <- paste0("Choose a ", label, ": ")
    selectInput(var, label, choices = choices, multiple = multiple)
}
```

This is a good idea if we saw that pattern repeated in multiple places -- but we probably also need to introduce some additional components for the ID, otherwise we risk generating controls with the same id. This is the problem of namespacing -- we want to have a hierarchy of names. This is one of the big problems that modules solve.

If we had a lot of variables, we could use functional programming to generate them. The key idea is to capture all the data needed to generate the columns in a single data frame. Dataframes are useful because they gemeralize to any number of arguments.

Here is an example: 

```{r}
vars <- tibble::tribble(~ var, ~label,
                        "t_type", "trafficking type",
                        "t_sub_type", "trafficking sub type",
                        "victim_gender", "gender")

# use pmap to turn each row in the df to a call to ngoSelectInput
vars %>%
    pmap(ngo_select_input) %>% # create a select input for each row 
    map(box_header) %>% # wrap each in a box header
    fluidRow(!!!.) # splat into a fluidRow 
```

we could be even more advanced and generate the call to dplyr::filter

```{r}
# create filter calls
(select <- map(vars$var, function(v) rlang::expr(.data[[!!v]] == input[[!!v]])))

# splat
filter(ngo, !!!select)
```

# Using Modules

functions are great but they are only useful for extracting pure ui code or pure computation used inside reactives. They don't help if we wish to build more complicated components that link ui and server.

A shiny module is a pair of functions corresponding to the front end UI and the backend server function. 

They essentially allow us to create an app within an app. They force isolation of behavior so there is no stepping on toes.

## Without Modules

As an example of why we need modules -- consider an app that has the user input a date as a string and validate it 

```{r}
library(lubridate)

ui <- fluidPage(
    textInput("date", "When were you born? (yyyy-mm-dd)"),
    textOutput("error"),
    textOutput("age"))

server <- function(input, output, session) {
    birthday <- reactive({
        req(input$date)
        ymd(input$date, quiet = TRUE)})

    age <- reactive({
        req(birthday())
        (birthday() %--% today() %/% years(1))})
    
    output$error <- renderText({
        if (is.na(birthday())) "Please enter valid date in yyyy-mm-dd form"})

    output$age <- renderText({paste0("You are ", age(), "years old")})
}

shinyApp(ui, server)
```

It is plausible that as the app gets bigger we might want to use the date control in multiple places -- so lets extract it into functions

```{r}
ymdInputUI <- function(label) {
    label <- paste0(label, " (yyyy-mm-dd)")

    fluidRow(
        textInput("date", label),
        textOutput("error"))}

ymdInputServer <- function(input, output, session) {
    date <- reactive({
        req(input$date)
        ymd(input$date, quiet = TRUE)})

    output$error <- renderText({
        if (is.na(date())) "Please enter valid date in yyyy-mm-dd form"})

    date
}

ui <- fluidPage(
    ymdInputUI("When were you born?"),
    textOutput("age"))

server <- function(input, output, session) {
    birthday <- ymdInputServer(input, output, session)

    age <- reactive({
        req(birthday())
        (birthday() %--% today() %/% years(1))})

    output$age <- renderText({
        paste0("You are ", age(), " years old")})}
```


There are two problems with this app: 

- it always assumes that the control is called date. This means we can't have two controls in the same app.
- the ui has an output with id error that you can't see from reading just the UI code. This makess it easy to break the app 

```{r}
ui <- fluidPage(
    ymdInputUI("When were you born?"),
    textOutput("error")
)
```

These problems arise because we've used functions to isolate local variables. The code is simpler to understand because we can't access the internals of the functions -- but there is another way that shiny code can interface: through the names of input and output controls.

This is the problem modules are made to solve. 

## Making a Module 

To convert the code above into a module, we need to make two changes: 

1. add an id argument to our UI component
2. use it with a special NS() function. This creates a namespace, or space of names, that is unique to the module 

```{r}
ymdInputUI <- function(id, label) {
    ns <- NS(id)
    label <- paste0(label, " (yyyy-mm-dd)")

    fluidRow(
        textInput(ns("date"), label),
        textOutput(ns("error")))}
```

The key idea is that the argument to NS() is supplied by the person using the component, and the args to the function it produces is supplied by the person who wrote the component. This ensures that the final name combines properties needed by both the app and module author. 

We now need to specify an id when creating the UI. This is important because it puts this id in the same place as all the others, so it is easy to spot if we've used the same input id in multiple places.

```{r}
ui <- fluidPage(
    ymdInputUI("birthday", "When were you born?"),
    textOutput("age")
)
```

Now we make a similar change to the server side of the module. Here, instead of NS, we use callModule which automatically tweaks the input and output so it looks for date inside the id namespace. Its doing something like input[[paste(id, "-", birthday)]]. 

```{r}
ymdInput <- function(id) {
    callModule(id = id, function(input, output, session) {
        date <- reactive({
            req(input$date)
            ymd(input$date, quiet = TRUE)})

        output$error <- renderText({
            if (is.na(date())) "Please enter valid date in yyyy-mm-dd form"})

        date
    })
}
```

Now the arguments to ymdInput have changed -- we pass in the id and shiny takes care of automatically plumbing up the input, output and session in the appropriate namespaced way. 

```{r}
server <- function(input, output, session) {
    birthday <- ymdInput("birthday")

    age <- reactive({
        req(birthday())
        (birthday() %--% today()) %/% years(1)})

    output$age <- renderText({
        paste0("You are ", age(), " years old")})}
```

## Limited Selection + Other

Consider the following app which provides a way to select gender that is sensitive to the many possible ways that people can express their gender 

```{r}
ui <- fluidPage(
    radioButtons("gender", "Gender:",
                 choiceValues = list("male", "female", "self-described", "na"),
                 choiceNames = list("Male", "Female",
                                textInput("gender_self", NULL,
                                          placeholder = "Self-Described"),
                                "Prefer not to say"),
                 selected = "na"),
    textOutput("txt"))

server <- function(input, output, session) {
    observeEvent(input$gender_self, {
        req(input$gender_self)
        updateRadioButtons(session, "gender",
                           selected = "self-described")})

    gender <- reactive({
        if (input$gender == "self-described") {
            input$gender_self
        } else {
            input$gender
        }})

    output$txt <- renderText({
        paste("You chose", gender())})}

sp()
```

Now, converting to a module and generalizing a little 

```{r}
radioButtonsWithOther <- function(id, label, choices,
                                  selected = NULL, placeholder = NULL) {
    ns <- NS(id)

    radioButtons(ns("primary"), "Gender:",
                 choiceValues = c(names(choices), "other"),
                 choiceNames = c(
                     unname(choices),
                     list(textInput(ns("other"),
                                    NULL,
                                    placeholder = NULL))),
                     selected = selected)
}

radioButtonsWithOtherServer <- function(input, output, session) {
    observeEvent(input$primary, {
        req(input$other)
        updateRadioButtons(session, "primary", selected = "other")})

    reactive({
        if (input$primary == "other") {
            input$other
        } else {
            input$primary
        }})}
```

```{r}
ui <- fluidPage(
    radioButtonsWithOther("gender",
                          choices = list(
                              male = "Male",
                              female = "Female",
                              na = "Prefer not to say"),
                          placeholder = "Self-Described",
                          selected = "na"),
    textOutput("txt"))

server <- function(input, output, session) {
    gender <- callModule(radioButtonsWithOtherServer, "gender")

    output$txt <- renderText(paste("You chose", gender()))
}

sp()
```

## Hierarchical Select Boxes

```{r}
country_df <- countrycode::codelist %>%
    as_tibble() %>%
    select(iso3c, continent, country = cow.name) %>%
    filter(!is.na(continent), !is.na(country))

continents <- country_df$continent %>% unique() %>% sort()

ui <- fluidPage(
    selectInput("continent", "Continent", choices = continents),
    selectInput("country", "Country", choices = NULL))

server <- function(input, output, session) {
    countries <- reactive({
        country_df[country_df$continent == input$continent, ,drop = FALSE]})

    observeEvent(input$continent, {
        updateSelectInput(session, "country", choice = countries()$country)})}

sp()
```

## Modal

```{r}
modalContinue <- function(id) {
    ns <- NS(id)

    showModal(
        modalDialog(
            "This is really important!",
            "Are you sure you want to continue?",
            footer = list(
                actionButton(ns("yes"), "yes"),
                actionButton(ns("no"), "no")),
            easyClose = FALSE))}

modalContinueServer <- function(id) {
    module <- function(input, output, session) {
        val <- reactiveVal(NULL)

        observeEvent(input$yes, {
            removeModal()
            val(TRUE)})

        observeEvent(input$no, {
            removeModal()
            val(FALSE)})

        reactive(val())}
    callModule(module, id)}
```

```{r}
ui <- fluidPage(
    actionButton("go", "Go"),
    textOutput("result"))

server <- function(input, output, session) {
    observeEvent(input$go, {
        modalContinue("ok")})

    result <- modalContinueServer("ok")

    output$result <- renderText(result())
}

sp()
```

# Exercises

1. The following app plots user selected variables from the msleep dataset for three different types of mammals (carnivores, omnivores, herbavores), with one tab for each type of mammal. Remove the redundancy in the selectInput definitions with the use of functions

```{r}

selectInputMammal <- function(id, label, selected = "sleep_rem") {
    ns <- NS(id)
    
    choices <- c("sleep_total", "sleep_rem", "sleep_cycle",
                 "awake", "brainwt", "bodywt")
    
    selectInput(inputId = ns("mammal"),
                label = label,
                choices = choices,
                selected = selected)
}

ui <- fluidPage(
    selectInputMammal("x", "X-axis"),
    selectInputMammal("y", "Y-axis", selected = "sleep_total"),
    tabsetPanel(id = "vore",
                tabPanel("Carnivore",
                         plotOutput("plot_carni")),
                tabPanel("Omnivore",
                         plotOutput("plot_omni")),
                tabPanel("Herbivore",
                         plotOutput("plot_herbi"))))
```

2. Further remove redundancy by modularizing how subsets and plots are created 

```{r}
selectMammal <- function(id) {
    ns <- NS(id)
    
    tagList(
        selectInput(ns("x"),
                    label = "x axis",
                    choices = c("sleep_total", "sleep_rem", "sleep_cycle", 
                                "awake", "brainwt", "bodywt"),
                    selected = "sleep_rem"),
        selectInput(ns("y"),
                    label = "y axis",
                    choices = c("sleep_total", "sleep_rem", "sleep_cycle", 
                                "awake", "brainwt", "bodywt"),
                    selected = "sleep_total"),
        tabsetPanel(id = ns("vore"),
                    tabPanel(title = "Carnivore",
                             value = "carni",
                             plotOutput(ns("plot_carni"))),
                    tabPanel(title = "Omnivore",
                             value = "omni",
                             plotOutput(ns("plot_omni"))),
                    tabPanel(title = "Herbivore",
                             value = "herbi",
                             plotOutput(ns("plot_herbi")))))
}

simple_plot <- function(data, x, y) {
    data %>%
        ggplot(aes_string(x, y)) +
        geom_point()
}

mammal_server <- function(input, output, session) {
    df <- reactive(msleep %>% filter(vore == input$vore))
    
    output$plot_carni <- renderPlot(simple_plot(df(), input$x, input$y))
    output$plot_omni <- renderPlot(simple_plot(df(), input$x, input$y))
    output$plot_herbi <- renderPlot(simple_plot(df(), input$x, input$y))
}

ui <- fluidPage(selectMammal(id = "mammals"))

server <- function(input, output, session) {
    callModule(module = mammal_server, id = "mammals")
}

sp()

```

### Extra Examples of Modules

This is from https://shiny.rstudio.com/articles/modules.html

A small but complete application demonstrating a simple "counter" module

```{r}
# module UI
counterButton <- function(id, label = "Counter") {
    # initialize namespace for ui. This returns input[[paste(id, "-", "counter")]]
    # This encapsulates the modules UI
    ns <- NS(id)
    
    tagList(
        actionButton(ns("button"), label = label),
        verbatimTextOutput(ns("out")))}

selectInputMammal <- function(id, label, selected = "sleep_rem") {
    ns <- NS(id)
    
    choices <- c("sleep_total", "sleep_rem", "sleep_cycle",
                 "awake", "brainwt", "bodywt")
    
    selectInput(inputId = ns("mammal"),
                label = label,
                choices = choices,
                selected = selected)
}

# module server side logic
counter <- function(input, output, session) {
    # initialize counter
    count <- reactiveVal(0)

    observeEvent(input$button, {
        count(count() + 1)
    })

    
    output$out <- renderText(count())

    # return a reactive value
    count
}

subset_and_plot_mammal <- function(input, output, session) {
    
}

ui <- fluidPage(counterButton("counter1", "Counter #1"))

server <- function(input, output, session) {
    # callmodule initializes the module 
    callModule(counter, "counter1")
}

sp()
```

Modules can represent input, output, or both. A module is composed of two functions that represent

1. a piece of UI 
2. a fragment of server logic that uses that UI 

### Creating UI 

The first argument to a UI should always be id. This is the namespace for the module. Note that the namespace for the module is decided by the caller at the time the module is used, not decided by the author at the time the module is writen. 

Here is an example for a CSV file input module

```{r}
csvFileInput <- function(id, label = "CSV File") {
    # create a namespace function using the provided id
    ns <- NS(id)

    tagList(
        fileInput(ns("file"), label),
        checkboxInput(ns("heading"), "Has heading"),
        selectInput(ns("quote"), "Quote",
                    c("None" = "",
                      "Double Quote" = "\"",
                      "Single Quote" = "'")))}
```

This looks similar to the ui.R file, with the following differences

1. The function body starts with `ns <- NS(id)`. This takes the string `id` and creates a namespace function 
2. Anything input or any output id needs to be wrapped in a call to `ns()`. 
3. The results are wrapped in tagList instead of fluidPage, pageWithSidebar, etc. You only need to use tagList if you want to return a UI fragment that consists of multiple UI objects; if we were just returning a div or some specific input, we could skip taglist 

Thanks to namespacing, we only need to make sure that the IDs "file", "heading", and "quote" are unique within this function, rather than unique across the entire app. 

### Server Functions

Module server functions should be named like their corresponding module ui functions, but without the input/output/ui suffix. Since our UI function was called csvFileInput, we'll call our server function csvFile

```{r}
csvFile <- function(input, output, session, stringsAsFactors) {
    # the selected file, if any
    userFile <- reactive({
        validate(need(input$file, message = FALSE))
        input$file})

    # the users data, parsed into a data frame
    dataframe <- reactive({
        read.csv(userFile()$datapath,
                 header = input$heading,
                 quote = input$quote,
                 stringsAsFactors = stringsAsFactors)})

    # we can run observers in here if we want to
    observe({
        msg <- sprintf("File %s was uploaded", userFile()$name)
        cat(msg, "\n")})

    return(dataframe)
}
```

This looks very similar to a server function. We can also add as many extra arguments to the function as desired, even ... if it makes sense to do so.

Inside the function input$file refers to ns("file") in the ui function. If the example had outputs, we could similarly match up ns("plot") with output$plot. 

The input, output, session objects are special in that they are scoped to the specific namespace that matches up with our ui function. Conversely, they can -not- access inputs/outputs outside of the namespace, nor can they directly access reactive expressions and reactive values from elsewhere in the application. Only lexically scoped reactive expressions/values can be used. 

If a module needs to use a reactive expression, take the reactive expression as a function parameter. If a module wants to return reactive expressions to the calling app, then return a list of reactive expressions from the function.

If a module needs to access an input that isn't part of the module, the containing app should pass the input value wrapped in a reactive expression (i.e. reactive(...)): `callModule(myModule, "myModule1", reactive(input$checkbox1))`

### Using Modules

This is how we would use the csvFileInput and csvFile functions together as a module in an app

```{r}
ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            csvFileInput("datafile", "User Data (.csv format)")),
        mainPanel(
            dataTableOutput("table"))))

server <- function(input, output, session) {
    # arg1 is the module server function
    # arg2 is the id we will use as a namespace. this needs to match the ui id
    datafile <- callModule(csvFile, "datafile", stringsAsFactors = FALSE)

    output$table <- renderDataTable(datafile())
}

shinyApp(ui, server)
```

The UI function csvFileInput is called directly using "datafile" as the id. In this case we insert the generated UI into the sidebar. 

The module server function is not called directly; instead, call the callModule function and provide the module server function as the first argument. 

### Output Example

Here is an example that consists of two linked scatterplots (selecting an area on one plot will highlight observations on both plots). 

First we made the module UI function. We want two plots, plot1 and plot2 side by side with a common brush id of brush. 

```{r}
linkedScatterUI <- function(id) {
    ns <- NS(id)

    fluidRow(
        column(6, plotOutput(ns("plot1"), brush = ns("brush"))),
        column(6, plotOutput(ns("plot2"), brush = ns("brush"))))}
```

The module server function is next. Besides the mandatory in, out, sess parameters, we need to know the dataframe to plot and the column names that should be used for the x and y axis' for each of the two plots.

To allow the dataframe and columns to change in response to user actions, the data, left, and right must all be reactive expressions.

```{r}
linkedScatter <- function(input, output, session,
                          data, left, right) {
    # yields the DF with an additional column selected_
    # that indicates whether that observation is brushed
    dataWithSelection <- reactive({
        brushedPoints(data(), input$brush, allRows = TRUE)})

    output$plot1 <- renderPlot({
        scatterPlot(dataWithSelection(), left())})

    output$plot2 <- renderPlot({
       scatterPlot(dataWithSelection(), right())})

    return(dataWithSelection)}
```

Notice that the linkedScatter function returns the dataWithSelection reactive. This allows the caller of the module to make use of the brushed data as well, such as showing it in a table below the plots for example.

For clarity, we can use the plotting code in a standalone function. 

```{r}
scatterPlot <- function(data, cols) {
    ggplot(data, aes_string(x = cols[1], y = cols[2])) +
        geom_point(aes(color = selected_)) +
        scale_color_manual(values = c("black", "#66D65C"), guide = FALSE)
}
```

```{r}
ui <- fixedPage(
    h2("Module Example"),
    linkedScatterUI("scatters"),
    textOutput("summary"))

server <- function(input, output, session) {
    df <- callModule(linkedScatter,
                     "scatters",
                     reactive(mpg),
                     left = reactive(c("cty", "hwy")),
                     right = reactive(c("drv", "hwy")))

    output$summary <- renderText({
        sprintf("%d observation(s) selected",
                nrow(filter(df(), selected_)))})}

sp()
```

### Nesting Modules

Modules can use other modules. When doing so, the outer module's UI function calls the inner module's ui function ensure that the id is wrapped in ns(). 

In the following example, outerUI called innerUI. Notice that the id argument is `ns("inner1")`

```{r}
innerUI <- function(id) {
    ns <- NS(id)
    "This is the inner UI"
}

outerUI <- function(id) {
    ns <- NS(id)
    wellPanel(
        innerUI(ns("inner1"))
    )
}
```

As for the module server functions, just ensure that the call to callModule for the inner module happens inside the outer module's server function. There is generally no need to use ns()

```{r}
inner <- function(in, out, sess) {
    # inner logic
}

outer <- function(in, out, sess) {
    innerResult <- callModule(inner, "inner1")
    # outer logic
}
```

### Using renderUI within modules

Inside of a module, we may want to use uiOutput / renderUI. If our renderUI block contains inputs/outputs, we need to use ns() to wrap our ID arguments just like the examples above. Unfortunately, those `ns` instances were created using NS(id), and in this case there is no `id` parameter to use. What do? 

The session parameter can provide the ns for you; just call `ns <- session$ns`. This will put the id in the same namespace as the session. 

```{r}
columnChooserUI <- function(id) {
    ns <- NS(id)
    uiOutput(ns("controls"))
}

columnChooser <- function(input, output, session, data) {
    output$controls <- renderUI({
        ns <- session$ns
        selectInput(ns("col"), "Columns", names(data), multiple = TRUE)})

    return(reactive({
        validate(need(input$col, FALSE))
        data[, input$col]}))
}
```

```{r}
mod_ui <- function(id, title) {
    ns <- NS(id)

    tagList(
        h2(title),
        selectInput(
            ns("what"),
            "What",
            choices = c("iris", "mtcars", "airquality")),
        plotOutput(ns("plot")))
}

mod_server <- function(input, output, session) {
    output$plot <- renderPlot({
        switch(
            input$what,
            "iris" = plot(iris),
            "mtcars" = plot(mtcars),
            "airquality" = plot(airquality))})
}

ui <- fluidPage(
    column(6, mod_ui("ui1", title = "left")),
    column(6, mod_ui("ui2", title = "right")))

server <- function(input, output, session) {
    callModule(mod_server, "ui1")
    callModule(mod_server, "ui2")
}

shinyApp(ui, server)
```

2.5 : rebuild the birthday app from earlier with modules

```{r}
library(lubridate)

ui <- fluidPage(
  textInput("date", "When were you born? (yyyy-mm-dd)"),
  textOutput("error"),
  textOutput("age")
)

server <- function(input, output, session) {
  birthday <- reactive({
    req(input$date)
    ymd(input$date, quiet = TRUE)
  })
  
  output$error <- renderText({
    if (is.na(birthday())) {
      "Please enter valid date in yyyy-mm-dd form"
    }
  })
  age <- reactive({
    req(birthday())
    (birthday() %--% today()) %/% years(1)
  })
  output$age <- renderText({
    paste0("You are ", age(), " years old")
  })
}

sp()
```


```{r}
birthdayUI <- function(id) {
    ns <- NS(id)

    tagList(textInput(ns("date"), "When were you born? (yyyy-mm-dd)"),
            textOutput(ns("error")),
            textOutput(ns("age")))
}

birthday <- function(input, output, session) {
    birthday <- reactive({
        req(input$date)
        ymd(input$date, quiet = TRUE)
    })

    age <- reactive({
        req(birthday())
        (birthday() %--% ymd(Sys.Date())) %/% years(1)
})
    output$error <- renderText({
        if (is.na(birthday())) {
            "Please enter a valid date (yyyy-mm-dd)"
        }
    })

    output$age <- renderText({
        paste0("you are ", age(), " years old")
    })
}

ui <- fluidPage(birthdayUI("datecheck"))

server <- function(input, output, session) {
    callModule(birthday, "datecheck")
}

sp()
```

```{r}
library(lubridate)

ymdInputUI <- function(id, label) {
    ns <- NS(id)
    label <- paste0(label, " (yyyy-mm-dd)")

    fluidRow(textInput(ns("date"), label),
             textOutput(ns("error")))}

ui <- fluidPage(
    ymdInputUI("birthday", "What were you born?"),
    textOutput("age"))

ymdInput <- function(id) {
    callModule(id = id, function(input, output, session) {
        date <- reactive({
            req(input$date)
            ymd(input$date, quiet = TRUE)})

        output$error <- renderText({
            if (is.na(date())) {
                "Please enter valid date in yyyy-mm-dd form"
            }
        })
        date
    })
}

server <- function(input, output, session) {
    birthday <- ymdInput("birthday")

    age <- reactive({
        req(birthday())
        (birthday() %--% today()) %/% years(1)
    })

    output$age <- renderText({
        paste0("You are ", age(), " years old")
    })
}

sp()
```

3. Maybe. If there are many interacting parts, then whenever one piece of that puzzle gets updated, others reactively get updated as well. If we modularize our code, then each of the pieces works independently of one another. This means that there is less computation spent on updating other components that are not the target. 

