# Dynamic UI 

```{r}
library(tidyverse)
library(shiny)

sp <- function() {
    shinyApp(ui, server)
}
```

There are three key techniques for creating dynamic user interfaces: 

- using update functions to modify the parameters of input controls 
- using tabsetPanel to conditionally show and hide parts of the ui 
- using uiOutput and renderUI to generate selected parts of the user interface with code 

A word of warning: These can make your app harder to reason about 

## Updating Inputs

We will start with a technique that allows the user to modify an input after it has been created. Every input control, i.e. textInput is paired with an update function, i.e. updateTextInput, that allows you to modify the control after it has been created. 

The update functions all take session as the first argument and the name of the input as the second argument. 

The key idea in the app below is that we can use observeEvent to trigger updateNumericInput whenever the min or max inputs change. 

```{r}
ui <- fluidPage(
    numericInput("min", "Minimum", 0),
    numericInput("max", "Maximum", 3),
    sliderInput("n", "n", min = 0, max = 3, value = 1))

server <- function(input, output, session) {
    observeEvent(input$min, updateNumericInput(session, "n", min = input$min))
    observeEvent(input$max, updateNumericInput(session, "n", max = input$max))
}

shinyApp(ui, server)
```

## Simple Uses

The simplest uses of update functions are to provide small conveniences for the user. 

For example, you may want to make it easy to reset parameters back to their initial value. 

```{r}
ui <- fluidPage(
    sliderInput("x1", "x1", 0, min = -10, max = 10),
    sliderInput("x2", "x2", 0, min = -10, max = 10),
    sliderInput("x3", "x3", 0, min = -10, max = 10),
    actionButton("reset", "Reset"))

server <- function(input, output, session) {
    observeEvent(input$reset, {
        updateNumericInput(session, "x1", value = 0)
        updateNumericInput(session, "x2", value = 0)
        updateNumericInput(session, "x3", value = 0)})}

sp()
```

A similar action is to tweak the text of an action button so that you know exactly what its going to do

```{r}
ui <- fluidPage(
    numericInput("n", "Simulations", 10),
    actionButton("simulate", "Simulate"))

server <- function(input, output, session) {
    observeEvent(input$n, {
        label <- paste0("Simulate ", input$n, " times")
        updateActionButton(session, "simulate", label = label)})}

sp()
```

## Hierarchical Select Boxes

A more complicated, but useful, application of the update function is to allow interactive drill down across multiple categories. 

```{r}
sales <- vroom::vroom("sales_data_sample.csv", col_types = list())
```

There is a natural hierarchy in the data:

- each territory contains customers 
- each customer has multiple orders
- each order contains rows

We want to create a user interface where you can:

- select a territory to see all customers 
- select a customer to see all orders 
- select an order to see the underlying rows 

The ui has three select boxes and one output table. The choices for customername and ordernumber select boxes are dynamically generated, so we set choices = NULL

```{r}
ui <- fluidPage(
    selectInput("territory", "Territory", choices = unique(sales$TERRITORY)),
    selectInput("customername", "Customer", choices = NULL),
    selectInput("ordernumber", "Order Number", choices = NULL),
    tableOutput("data"))
```

In the server function, we can work top down:

- create a reactive territory() that contains the rows from sales that match the selected territory
- whenever territory changes, update the list of choices in the input$customername select box 
- create another reactive, customer(), which contains rows from territory() that match the selected customer
- whenever customer changes, update the list of choices in the input$ordernumber select box 
- display the selected orders in output$data 

```{r}
server <- function(input, output, session) {
    territory <- reactive(sales %>% filter(TERRITORY == input$territory))
    
    observeEvent(territory(), {
        choices <- unique(territory()$CUSTOMERNAME)
        updateSelectInput(session, "customername", choices = choices)})

    customer <- reactive({
        req(input$customername)
        territory() %>% filter(CUSTOMERNAME == input$customername)})
    
    observeEvent(customer(), {
        choices <- unique(customer()$ORDERNUMBER)
        updateSelectInput(session, "ordernumber", choices = choices)})

    output$data <- renderTable({
        req(input$ordernumber)
        customer() %>%
            filter(ORDERNUMBER == input$ordernumber) %>%
            select(QUANTITYORDERED, PRICEEACH, PRODUCTCODE)})}

sp()
```

## Circular References

From shinys perspective, using an update function to modify value is no different than a user modifying the value by clicking or typing. This can lead to hairy situations in which we get feedback loops. Here is an example:

```{r}
ui <- fluidPage(numericInput("n", "n", 0))

server <- function(input, output, session) {
    observeEvent(input$n,
                 updateNumericInput(session, "n", value = input$n + 1))}

sp()
```

This problem can crop up if you update multiple controls that depend on one another, as in the next example

## Inter-related Inputs

One place where its easy to end up with circular references is when you have multiple sources of truth in an app. 

```{r}
ui <- fluidPage(
    numericInput("temp_c", "Celsius", NA),
    numericInput("temp_f", "Fahrenheit", NA))

server <- function(input, output, session) {
    observeEvent(input$temp_f, {
        c <- round((input$temp_f - 32) * 5 / 9)
        updateNumericInput(session, "temp_c", value = c)})

    observeEvent(input$temp_c, {
        f <- round((input$temp_c * 9 / 5) + 32)
        updateNumericInput(session, "temp_f", value = f)})
}

sp()
```

Sometimes the app triggers multiple changes. There is no way around this problem because we have one idea (temperature) with multiple expressions in the app (celsius and fahrenheit). Here we are fortunate that the cycle quickly converges to a value that satisfies both constraints. In general we are best off avoiding this problem. 

## Exercises 

1. 

```{r}
ui <- fluidPage(
    numericInput("year", "year", value = 2020),
    dateInput("date", "date"))

server <- function(input, output, session) {
    observeEvent(input$year, {
        newDateMin <- reactive(format(paste0(input$year, "-01-01")))
        newDateMax <- reactive(format(paste0(input$year, "-12-31")))
        updateDateInput(session, "date",
                        min = newDateMin(),
                        max = newDateMax())})
}

sp()
```

2. Complete the interface below with a server function that updates input$county choices based on input$state. For an added challenge, also change the label from "County" to "Parrish" for Louisiana and "Borrough" for "Alaska".

```{r}
library(openintro)
states <- unique(county$state)

county %>% filter(state == "Massachusetts") %>% select(name) %>% map(as.character) %>% flatten_chr()

ui <- fluidPage(
    selectInput("state", "State", choices = states),
    selectInput("county", "County", choices = NULL))

server <- function(input, output, session) {
    statein <- reactive({
        chosen_state <- reactive(county %>%
                                 filter(state := input$state) %>%
                                 select(name) %>%
                                 map(as.character) %>%
                                 flatten_chr())})

    observeEvent(statein(), {
        updateSelectInput(session, "county", choices = statein())})
}

sp()
```

FIX THIS^^^

3. Complete the user interface below with a server function that updates input$country choices based on the input$continent. Use output$data to display all matching rows 

```{r}
library(gapminder)
continents <- unique(gapminder$continent)

ui <- fluidPage(
    selectInput("continent", "Continent", choices = continents),
    selectInput("country", "Country", choices = NULL),
    tableOutput("data"))

server <- function(input, output, session) {
    countries <- reactive({
        gapminder %>% filter(continent == input$continent)})

    observeEvent(countries(), {
        choices <- countries()$country %>% unique()
        updateSelectInput(session, "country", choices = choices)})

    output$data <- renderTable({
        req(input$country)
        countries() %>%
            filter(country == input$country)})
}

sp()
```

4. Extend the previous app so that you can also choose to select no continent and see all the countries. 

```{r}
continents <- unique(c("all", as.character(gapminder$continent)))

ui <- fluidPage(
    selectInput("continent", "Continent", choices = continents),
    selectInput("country", "Country", choices = NULL),
    tableOutput("data"))

server <- function(input, output, session) {
    countries <- reactive({
        if (input$continent == "all") {
            gapminder
        } else {
            gapminder %>% filter(continent == input$continent)
        }})
        ## gapminder %>% filter(continent == input$continent)})
        
    observeEvent(countries(), {
        choices <- countries()$country %>% unique()
        updateSelectInput(session, "country", choices = choices)})

    output$data <- renderTable({
        req(input$country)
        countries() %>%
            filter(country == input$country)})
}

sp()
```

5. Linear dependent inputs

# Dynamic Visibility

The next step up in complexity is selectively showing and hiding parts of the ui. 

First up we will look at concealing optional ui in a tabset. This allows us to show and hide the ui as needed without having to regenerate it from scratch. Currently we need a smidgen of css to make the technique work:

tag$style("#switcher { display:none; }") where #switcher is the #id of your tabset panel 

```{r}
ui <- fluidPage(
    tags$style("#switcher { display:none; }"),
    sidebarLayout(
        sidebarPanel(
            selectInput("controller", "Show", choices = paste0("panel", 1:3))),
        mainPanel(
            tabsetPanel(
                id = "switcher",
                tabPanel("panel1", "Panel 1 Content"),
                tabPanel("panel2", "Panel 2 Content"),
                tabPanel("panel3", "Panel 3 Content")))))

server <- function(input, output, session) {
    observeEvent(input$controller, {
        updateTabsetPanel(session, "switcher", selected = input$controller)})}

sp()
```

## Conditional UI

Imagine that you want an app that allows the user to simulate from the normal, uniform, and exponential distributions. Here we put the unique user interface for each distribution in its own tabPanel and then arrange the three tabs into a tabsetPanel

```{r}
# this will be embedded into ui
parameter_tabs <- tagList(
    tags$style("#params { display:none; }"),
    tabsetPanel(id = "params",
                tabPanel("normal",
                         numericInput("mean", "mean", value = 1),
                         numericInput("sd", "standard deviation",
                                      min = 0, value = 1)),
                tabPanel("uniform",
                         numericInput("min", "min", value = 0),
                         numericInput("max", "max", value = 1)),
                tabPanel("exponential",
                         numericInput("rate", "rate", value = 1, min = 0))))

ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            selectInput("dist", "Distribution",
                        # match choices to input$dist
                        choices = c("normal", "uniform", "exponential")),
            numericInput("n", "Number of Samples", value = 100),
            parameter_tabs),
        mainPanel(plotOutput("hist"))))

server <- function(input, output, session) {
    observeEvent(input$dist, {
        updateTabsetPanel(session, "params",
                          selected = input$dist)})

    sample <- reactive({
        switch(input$dist,
               normal = rnorm(input$n, input$mean, input$sd),
               uniform = runif(input$n, input$min, input$max),
               exponential = rexp(input$n, input$rate))})

    output$hist <- renderPlot(hist(sample()))}

sp()
```

## Wizard Interface

We can use this idea to make a wizard, a type of interface that makes it easy to collect a bunch of information by spreading it across multiple web pages. Here we embed action buttons within each page making it easy to go forward and back. 

```{r}
ui <- fluidPage(
    tags$style("#wizard { display:none; }"),
    tabsetPanel(id = "wizard",
                tabPanel("page1",
                         "Welcome!",
                         actionButton("page12", "next")),
                tabPanel("page2",
                         "Only one page to go",
                         actionButton("page21", "prev"),
                         actionButton("page23", "next")),
                tabPanel("page3",
                         "You're Done!",
                         actionButton("page32", "prev"))))

server <- function(input, output, session) {
    switch_tab <- function(page) {
        updateTabsetPanel(session, "wizard", selected = page)}

    observeEvent(input$page12, switch_tab("page2"))
    observeEvent(input$page21, switch_tab("page1"))
    observeEvent(input$page23, switch_tab("page3"))
    observeEvent(input$page32, switch_tab("page2"))
}

sp()
```

# Dialog Boxes

Not yet implemented in the book. Check back!

# Creating UI with Code

There is one last technique this chapter covers that gives you the ability to create any controls (both inputs and outputs) with code in the server function. This technique allows us to regenerate the ui code while the app is running.

The technique has two components:

- use uiOutput to insert a placeholder in the user interface. This is run when the app launches and it leaves a hole that the server code can later fill in
- renderUI fills the placeholder with the ui generated in the server function

Here is a simple example that dynamically creates a different type of input control depending on an input. 

```{r}
ui <- fluidPage(
    textInput("label", "label"),
    selectInput("type", "type", c("slider", "numeric")),
    uiOutput("numeric"))

server <- function(input, output, session) {
    output$numeric <- renderUI({
        if (input$type == "slider") {
            sliderInput("dynamic", input$label, value = 0, min = 0, max = 10)
        } else {
            numericInput("dynamic", input$label, value = 0, min = 0, max = 10)}})}

sp()
```

Theres a problem: When you change controls, you lose the currently selected values. It is generally better practice to selectively show and hide ui as above. We can fix the problem by setting the value of the new input to the current value of the existing control 

```{r}
server <- function(input, output, session) {
    output$numeric <- renderUI({
        if (input$type == "slider") {
            sliderInput("dynamic", input$label,
                        value = isolate(input$dynamic),
                        min = 0, max = 10)
        } else {
            numericInput("dynamic", input$label,
                         value = isolate(input$dynamic),
                         min = 0, max = 10)}})}

sp()
```

isolate() ensures that we don't create a reactive dependency that would mean the code is rerun every time input$dynamic changes. We only want to change it when input$type or input$label changes. 

Dynamic UI is most useful when you are generating an arbitrary number of controls. This is a good use case for purrr 

## Multiple Controls 

An important use of renderUI() is when we need to create an arbitrary number of controls. 

The ui has a numericInput that controls the number of inputs, a uiOutput where the generated text boxes will go, and a textOutput that demonstrates that we have plumbed everything together correctly.

```{r}
ui <- fluidPage(
    numericInput("n", "Number of Colors",
                 value = 5, min = 1),
    uiOutput("col"),
    textOutput("palette"))
```

There are three key ideas in the server function:

- col_names() generated a character vector giving the name of each of the color inputs about to be generated 
- create the text boxes using map to create one textInput for input in col_names. output col inserts these text boxes in the ui placeholder created earlier 
- To generate the output we need to use a different trick. Usually we use $, but in this case, we are using a character vector. Instead we use [] and map_chr collects all values into a character vector and displays it in output\$palette

```{r}
server <- function(input, output, session) {
    col_names <- reactive(paste0("col", seq_len(input$n)))

    output$col <- renderUI(map(col_names(), ~ textInput(.x, NULL)))

    output$palette <- renderText(map_chr(col_names(), ~ input[[.x]]))
}

sp()
```

This app has a problem in that it deletes all the current values when you increase the number of colors. To fix this we need to use isolate again. 

Here is a more fleshed out version of the app

```{r}
ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            numericInput("n", "Number of Colors",
                         value = 5, min = 1),
            uiOutput("col")),
        mainPanel(
            plotOutput("plot"))))

server <- function(input, output, session) {
    col_names <- reactive(paste0("col", seq_len(input$n)))

    output$col <- renderUI({
        map(col_names(),
            ~ textInput(.x, NULL, value = isolate(input[[.x]])) %||% "")})

    output$plot <- renderPlot({
        cols <- map_chr(col_names(), ~ input[[.x]])
        cols[cols == ""] <- NA

        barplot(
            rep(1, length(cols)),
            col = cols,
            space = 0,
            axes = FALSE)})}

sp()
```

## Dynamic Filtering 

Now we are going to make an app that allows the user to dynamically filter any data frame. Each numeric input will get a range slider and each factor input will get a multi select. 

Lets start with a function that creates the UI for a single variable. It will return a slider for numeric inputs, a multi select for factor inputs and NULL for all other types 

```{r}
make_ui <- function(x, var) {
    if (is.numeric(x)) {
        rng <- range(x, na.rm = TRUE)
        sliderInput(var, var,
                    min = rng[1], max = rng[2], value = rng)
    } else if (is.factor(x)) {
        levs <- levels(x)
        selectInput(var, var,
                    choices = levs, selected = levs, multiple = TRUE)
    } else {
        # not supported
        NULL}}
```

The server function takes the variable and value of the input control and returns a logical vector saying whether or not to include each observation. The logical vector makes it easy to combine the results from multiple columns 

```{r}
filter_var <- function(x, var) {
    if (is.numeric(x)) {
        !is.na(x) & x >= var[1] & x <= var[2]
    } else if (is.factor(x)) {
        x %in% var
    } else {
        # no control, dont filter
        TRUE}}
```

Now we can use these functions to generate a simple filtering UI for the iris dataset.

```{r}
ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            make_ui(iris$Sepal.Length, "Sepal.Length"),
            make_ui(iris$Sepal.Width, "Sepal.Width"),
            make_ui(iris$Species, "Species")),
        mainPanel(tableOutput("data"))))

server <- function(input, output, session) {
    selected <- reactive({
        filter_var(iris$Sepal.Length, input$Sepal.Length) &
            filter_var(iris$Sepal.Width, input$Sepal.Width) &
            filter_var(iris$Species, input$Species)})

    output$data <- renderTable(head(iris[selected(), ], 12))}

sp()
```

The app had a lot of copy pasta, so it only got three columns. We can increase this with functional programming.

- in ui use map() to generate one control for each variable 
- in server(), use map() to generate the selection vector for each variable. Then use reduce to take the logical vector for each variable and combine into a single logical vector by &ing each vector together 

```{r}
ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            map(names(iris), ~ make_ui(iris[[.x]], .x))),
        mainPanel(tableOutput("data"))))

server <- function(input, output, session) {
    selected <- reactive({
        each_var <- map(names(iris), ~ filter_var(iris[[.x]], input[[.x]]))
        reduce(each_var, `&`)})

    output$data <- renderTable(head(iris[selected(), ], 12))}

sp()
```

From there its an easy generalization to work with any data frame. Here we illustrate it using the data frames in the datasets package, but this could easily be extended to uploaded data. 

```{r}
dfs <- keep(ls("package:datasets"), ~ is.data.frame(get(.x, "package:datasets")))

ui <- fluidPage(
    sidebarLayout(
        sidebarPanel(
            selectInput("dataset", label = "Dataset", choices = dfs),
            uiOutput("filter")),
        mainPanel(tableOutput("data"))))

server <- function(input, output, session) {
    data <- reactive({
        get(input$dataset, "package:datasets")})

    vars <- reactive(names(data()))

    output$filter <- renderUI(
        map(vars(), ~ make_ui(data()[[.x]], .x)))

    selected <- reactive({
        each_var <- map(vars(), ~ filter_var(data()[[.x]], input[[.x]]))
        reduce(each_var, `&`)})

    output$data <- renderTable(head(data()[selected(), ], 12))}

sp()
```

# Exercises

1. Given the simple app below

```{r}
ui <- fluidPage(
    selectInput("type", "type", c("slider", "numeric")),
    uiOutput("numeric"))

server <- function(input, output, session) {
    output$numeric <- renderUI({
        if (input$type == "slider") {
            sliderInput("n", "n", value = 0, min = 0, max = 100)
        } else {
            numericInput("n", "n", value = 0, min = 0, max = 100)}})}

sp()
```

How could we instead implement it using dynamic visibility? If we implement dynamic visibility, how could we keep the values in sync when the controls change? 

```{r}
dyn_widget <- function(x) {
    if (x == "slider") {
        sliderInput("dynamic", "n",
                    value = 1,
                    ## value = isolate(input$dynamic),
                    min = 0, max = 100)
    } else {
        numericInput("dynamic", "n",
                     value = 1,
                     ## value = isolate(input$dynamic),
                     min = 0, max = 100)}}

ui <- fluidPage(
    selectInput("type", "type", c("slider", "numeric")),
    uiOutput("numeric"))

server <- function(input, output, session) {
    output$numeric <- renderUI(dyn_widget(input$type))
}

sp()
```

```{r}
ui <- fluidPage(
    tags$style("#switchControl { display:none; }"),
    sidebarLayout(
        sidebarPanel(
            selectInput("controller", "Type", choices = c("slider", "numeric"))),
        mainPanel(
            tabsetPanel(
                id = "switchControl",
                tabPanel("slider",
                         dyn_widget("slider")),
                tabPanel("numeric",
                         dyn_widget("numeric"))))))

# FIX THIS

## server <- function(input, output, session) {
##     observeEvent(input$controller, {
##         updateTabsetPanel(session, "switchControl", selected = input$controller)
##         observeEvent(input$switchControl, {
##             val <- reactive(input$dynamic)
##             updateSelectInput(value = 5)
##             updateNumericInput(value = 6)
##     })})}

sp()
```
