# Your First Shiny App
<!-- https://mastering-shiny.org/basic-app.html -->

## A Minimal App

```{r}
library(shiny)

ui <- fluidPage(
    "hello, world!"
)

server <- function(input, output, session) {
    
}

shinyApp(ui, server)
```

## Adding UI Controls

This app shows all the built in data frames included in the datasets package. 

```{r}
ui <- fluidPage(
    selectInput("dataset", label = "Dataset", choices = ls("package:datasets")),
    verbatimTextOutput("summary"),
    tableOutput("table")
)

# tell shiny how to fill in the summary and table outputs
server <- function(input, output, session) {
    output$summary <- renderPrint({
        dataset <- get(input$dataset, "package:datasets")
        summary(dataset)
    })

    output$table <- renderTable({
        (dataset <- get(input$dataset, "package:datasets"))
    })
}

shinyApp(ui, server)
```

- fluidpage is a layout function that sets up the basic visual structure of the page. 

- selectInput is an input control that lets the user select from a box 

- verbatimtextoutput and tableoutput are output controls that tell shiny where to put rendered output. 

- almost every output written in shiny will follow this same pattern 

```{r, eval = FALSE}
output$id <- renderType({
    # expression that generates output rendertype expects
})
```

## Reducing Duplication with Reactive Expressions 

Generally when we wish to reduce duplication in R programming we either capture a value using a variable or capture a process using a function. Unfortunately, neither of these approaches work to reduce duplication in a shiny context. For that, we need a new mechanism: *reactive expressions*. 

We can create a reactive expression by wrapping a block of code in `reactive({...})` and assigning it to a variable, and we use the reactive  expression by calling it like a function. A reactive function is different than a function in that it only runs the first time it is called and then it caches the result until it needs to be updated. 

The app below behaves identically to the one above, but it works a little more efficiently because it only needs to retrieve the dataset once, not twice. 

```{r}
server <- function(input, output, session) {
    dataset <- reactive({
        get(input$dataset, "package:datasets")
    })

    output$summary <- renderPrint({
        summary(dataset())
    })

    output$table <- renderTable({
        dataset()
    })
}
```

## Exercises 

1. Create an app that greets the user by name. 

```{r}
ui <- fluidPage(
    textInput("name", "What's your name?"),
    textOutput("response")
)

server <- function(input, output, session) {
    output$response <- renderText({
        paste0("Hello, ", input$name)
    })
}

shinyApp(ui, server)
```

2. 
x was not properly defined in the context of input within the ui function. 

```{r}
ui <- fluidPage(
    sliderInput("x", "if x is", 1, 50, 30, 1),
    textOutput("by_five")
)

server <- function(input, output, session) {
    output$by_five <- renderText(
        paste0("\nthen x multiplied by 5 is\n", input$x * 5))
}

shinyApp(ui, server)
```

3. 

```{r}
ui <- fluidPage(
    sliderInput("x", "if x is", 1, 50, 30, 1),
    sliderInput("y", "and y is", 1, 50, 30, 1),
    textOutput("by_five")
)

server <- function(input, output, session) {
    output$by_five <- renderText({
        paste0("\nthen x multiplied by y is: ", input$x * input$y)))
}

shinyApp(ui, server)
```

4. 

```{r}
ui <- fluidPage(
    sliderInput("x", "if x is", 1, 50, 30, 1),
    sliderInput("y", "and y is", 1, 50, 5, 1),
    "then, (x * y) is", textOutput("product"),
    "and, (x * y) + 5 is", textOutput("product_plus5"),
    "and, (x * y) + 10 is", textOutput("product_plus10"),
)

server <- function(input, output, session) {
    output$product <- renderText({
        (product <- input$x * input$y)
    })

    output$product_plus5 <- renderText({
        (product <- input$x * input$y + 5)
    })

    output$product_plus10 <- renderText({
        (product <- input$x * input$y + 10)
    })
}

shinyApp(ui, server)
```

Then reduce using reactive 

```{r}
ui <- fluidPage(
    sliderInput("x", "if x is", 1, 50, 30, 1),
    sliderInput("y", "and y is", 1, 50, 5, 1),
    "then, (x * y) is", textOutput("product"),
    "and, (x * y) + 5 is", textOutput("product_plus5"),
    "and, (x * y) + 10 is", textOutput("product_plus10"),
)

server <- function(input, output, session) {
    product <- reactive({
        input$x * input$y
    })

    output$product <- renderText({product()})
    output$product_plus5 <- renderText({product() + 5})
    output$product_plus10 <- renderText({product() + 10})
}

shinyApp(ui, server)
```

5. 

```{r}
datasets <- data(package = "ggplot2")$results[, "Item"]

ui <- fluidPage(
    selectInput("dataset", "Dataset", choices = datasets),
    verbatimTextOutput("summary"),
    plotOutput("plot", 500, 500)
)

server <- function(input, output, session) {
    dataset <- reactive({
        get(input$dataset, "package:ggplot2")
    })

    output$summary <- renderPrint({
        dataset() %>% summary()
    })

    output$plot <- renderPlot({
        plot(dataset()[, 1:2])
    })
}

shinyApp(ui, server)
```

