# Tidy Evaluation

```{r}
library(tidyverse)
library(shiny)

sp <- function() {
    shinyApp(ui, server)
}
```

### Side Tangent : Using ggplot2 in packages

- if you know the mapping or facet specification is col in advance, use `aes(.data$col)` or `vars(.data$col)`
- if col is a variable that contains the column name as a character vector, use `aes(.data[[col]])` or `vars(.data[[col]])`
- if you want the behavior of col to look and feel like it would within `aes()` and `vars()`, use `aes({{col}})` or `vars({{col}})`

# Motivation

The key idea of tidy evaluation is that it blurs the line between two meanings of "variable":

- environment variables (env-vars for short) are programming variables. Formally, they are between names and values that are usually created by `<-` 
- data frame variables (data-vars for short) are statistical variables that live inside a data frame. In base R we usually access them with `$` and `[[`

```{r}
df <- data.frame(x = runif(3),
                 y = runif(3))

df$x
```

`df` is an env-var that contains two data-vars, `x` and `y`. 

In base R, we write code that looks like this

```{r}
diamonds[diamonds$x == 0 | diamonds$y == 0, ]
```

the dplyr equivalent uses tidy evaluation to allow us to refer to a data var as if it were an environment variable

```{r}
diamonds %>% filter(x == 0 | y == 0)
```

We usually use these verbs purely with data vars, but they work equally well with env-vars

```{r}
min_carat <- 1

diamonds %>% filter(carat > min_carat)
```

# Solutions

## Tidy Eval in Shiny Apps

The key to resolving this ambiguity is to make use of two pronouns that are built into tidy evaluation: `.data` and `.env`. These allow us to remove the ambiguity introduced by tidy evaluation. 

For example, we could rewrite the filter used above

```{r}
diamonds %>% filter(.data$carat > .env$min_carat)
```

This isn't immediately useful in shiny apps because the results from inputs are usually strings and using `.data$var` isn't going to work because it's going to look for a data-var called `var`, not a data-var stored in the env-var `var`. Fortunately, base R has a solution for this: `.data[[var]]`

```{r}
ui <- fluidPage(
    selectInput("var", "Variable", choices = names(diamonds)),
    tableOutput("output"))

server <- function(input, output, session) {
    data <- reactive(filter(diamonds, input$var > 0))
    output$output <- renderTable(head(data()))
}

sp()
```

This doesn't work because `input$var` isn't a data-var; it's an env-var containing the name of a data-var stored as a string. Unfortunately, it also fails to give a useful error message because `input$var` will be a string like "carat" and 

```{r}
"carat" > 0
```

This problem can be fixed by using `.data` and `[[`

```{r}
server <- function(input, output, session) {
    data <- reactive(filter(diamonds, .data[[input$var]] > 0))
    output$output <- renderTable(head(data()))
}

sp()
```

## Tidy Evaluation in Functions

We could use `.data` and `[[`, but it doesn't create very user friendly functions:

```{r}
filter_var <- function(df, var, val) {
    filter(df, .data[[var]] > val)
}

filter_var(diamonds, "carat", 4)
```

This is weird because it takes the name of the variable as a string, so it doesn't work in most other tidyverse functions. 

```{r}
filter_var <- function(df, var, val) {
    filter(df, {{var}} > val)
}

filter_var(diamonds, carat, 4)
```

## Parse

paste + parse + eval is tempting, but it has a major downside. It allows the users of your app to run arbitrary R code. This isn't that bad if only you are using the app, but it can create security holes if shared more widely. 

# Case Studies

## Plotting 

```{r}
ui <- fluidPage(
    selectInput("x", "X variable",
                choices = names(iris)),
    selectInput("y", "Y variable",
                    choices = names(iris)),
    plotOutput("plot"))

server <- function(input, output, session) {
    output$plot <- renderPlot({
        ggplot(iris, aes(.data[[input$x]], .data[[input$y]])) +
            geom_point(position = ggforce::position_auto()) +
            labs(x = input$x, y = input$y)})}

sp()
```

ggforce::position_auto was used to automatically spread the points out when one axis is discrete. Once tha basics of tidy evaluation are achieved, the challenge becomes how to make the app general enough to work with many different types of variable. 

Instead of position_auto, we could allow the user to pick the geom

```{r}
ui <- fluidPage(
    selectInput("x", "X variable", choices = names(iris)),
    selectInput("y", "Y variable", choices = names(iris)),
    selectInput("geom", "geom", c("point", "smooth", "jitter")),
    plotOutput("plot"))

server <- function(input, output, server) {
    plot_geom <- reactive({
        switch(input$geom,
               point = geom_point(),
               smooth = geom_smooth(se = FALSE),
               jitter = geom_jitter())})

    output$plot <- renderPlot({
        ggplot(iris, aes(.data[[input$x]], .data[[input$y]])) +
            plot_geom() +
            labs(x = input$x, y = input$y)})}

sp()
```

## Filtering and Arranging

```{r}
ui <- fluidPage(
    selectInput("var", "Select Variable", choices = names(mtcars)),
    sliderInput("min", "Minimum Value", 0, min = 0, max = 100),
    selectInput("sort", "Sort By", choices = names(mtcars)),
    tableOutput("data"))

server <- function(input, output, session) {
    observeEvent(input$var, {
        rng <- range(mtcars[[input$var]])
        updateSliderInput(session, "min",
                          value = rng[[1]],
                          min = rng[[1]],
                          max = rng[[2]])})

    output$data <- renderTable({
        mtcars %>%
            filter(.data[[input$var]] > input$min) %>%
            arrange(.data[[input$sort]])})}

sp()
```

Most other problems can be solved by combining this technique with existing skills. What if we wanted to conditionally sort in either ascending or descending order?

```{r}
ui <- fluidPage(
    selectInput("var", "Sort By", choices = names(mtcars)),
    checkboxInput("desc", "Descending Order?"),
    tableOutput("data"))

server <- function(input, output, session) {
    sorted <- reactive({
        if (input$desc) {
            arrange(mtcars, desc(.data[[input$var]]))
        } else {
            arrange(mtcars, .data[[input$var]])
        }
    })

    output$data <- renderTable(sorted())}

sp()
```

# Additional Challenges

## Selection Semantics

Most tidyverse functions (like mutate, filter, group_by, aes) have action semantics -- they can perform any action inside of them. Other functions have selection semantics. Instead of general computation we can select variables using a special domain specific language that includes helpers like starts\_with and ends\_with. The most important function that has selection semantics is select, but the set also includes many tidyr functions like pivot\_longer, pivot\_wider, separate, extract, and unite functions. 

Working with functions that use selection semantics is slihtly different than those that use action semantics because there is no .data pronoun. Instead we use the helper one\_of and all\_of/

```{r}
ui <- fluidPage(
    selectInput("vars", "Variables", names(mtcars), multiple = TRUE),
    tableOutput("data"))

server <- function(input, output, session) {
    output$data <- renderTable({
        req(input$vars)
        mtcars %>% select(one_of(input$vars))})}

sp()
```

if we wanted all of the variables except those selected, we could use `-one_of(input$vars)`

```{r}
ui <- fluidPage(
    selectInput("vars", "Variables", names(mtcars), multiple = TRUE),
    tableOutput("data"))

server <- function(input, output, session) {
    output$data <- renderTable({
        req(input$vars)
        mtcars %>% select(-one_of(input$vars))})}

sp()
```

## Multiple Variables

Working with multiple variables is trivial when you're working with a function that uses selection semantics -- we can just pass a character vector of variable names in to one\_of / all\_of. 

The challenge is operating on multiple variables when the function has action semantics, as is common with dplyr functions. 

There are two ways to work with multiple variables, depending on what version of dplyr you are working with. We can illustrate them with an app that allows you to select any number of variables to count their unique values.

```{r}
ui <- fluidPage(
    selectInput("vars", "Variables", names(mtcars), multiple = TRUE),
    tableOutput("count"))
```

In dplyr 0.8 and earlier, every function that uses action semantics also has a variant that has selection semantics with the suffix _at. The easiest approach is just to switch from action to selection semantics by changing the function.

```{r}
server <- function(input, output, session) {
    output$count <- renderTable({
        req(input$vars)

        mtcars %>%
            group_by_at(input$vars) %>%
            summarize(n = n())})}

sp()
```

dplyr 1.0 provides a more flexible approach: inside of any function with action semantics, we can use across() to access selection semantics

```{r}
server <- function(input, output, session) {
    output$count <- renderTable({
        req(input$vars)

        mtcars %>%
            group_by(across(all_of(input$vars))) %>%
            summarize(n = n())})}

sp()
```

Things are mildly more complicated for mutate and summarize because we also need to supply a function to perform the operation.

```{r}
ui <- fluidPage(
    selectInput("vars_g", "Group By", names(mtcars), multiple = TRUE),
    selectInput("vars_s", "Summarize", names(mtcars), multiple = TRUE),
    tableOutput("data"))

# dplyr 0.8
server <- function(input, output, session) {
    output$data <- renderTable({
        mtcars %>%
            group_by_at(input$vars_g) %>%
            summarize_at(.vars = input$vars_s, .funs = mean)})}

sp()

# dplyr 1.0
server <- function(input, output, session) {
    output$data <- renderTable({
        mtcars %>%
            group_by(across(all_of(input$vars_g))) %>%
            summarize(across(all_of(inputs$vars_s), mean), n = n())})}

sp()
```

## Action Semantics and User Supplied Data

There is one additional complication when working with user supplied data and action semantics. Consider the following app: it allows the user to upload a tsv file, then select a variable, then filter by it. 

```{r}
ui <- fluidPage(
    fileInput("data", "dataset", accept = ".tsv"),
    selectInput("var", "var", character()),
    numericInput("min", "min", 1, min = 0, step = 1),
    tableOutput("output"))

server <- function(input, output, session) {
    data <- reactive({
        req(input$data)
        vroom::vroom(input$data$datapath)})

    observeEvent(data(), {
        updateSelectInput(session, "var", choices = names(data()))})

    observeEvent(input$var, {
        val <- data()[[input$var]]
        updateNumericInput(session, "min", value = min(val))})

    output$output <- renderTable({
        req(input$var)

        data() %>%
            filter(.data[[input$var]] > input$min) %>%
            arrange(.data[[input$var]]) %>%
            head(10)})}

sp()
```

There is a subtle problem with the use of filter(). 

```{r}
df <- data.frame(x = 1, y = 2)

input <- list(var = "x", min = 0)

df %>% filter(.data[[input$var]] > input$min)
```

What happens when there is a data frame that contains a variable named input? 

```{r}
df <- data.frame(x = 1, y = 2, input = 3)

df %>% filter(.data[[input$var]] > input$min)
```

There is an error message because filter is trying to evaluate df$input$min.

This problem is due to the ambiguity of data variables and env variables. Tidy eval always prefers to use a data variable if both are available. We can resolve the ambiguity by telling filter not to look in the data frame for input and instead only use an env-variable.

```{r}
df

df %>% filter(.data[[input$var]] > .env$input$min)
```

Instead of .env, we could also use !!, e.g.

```{r}
df %>% filter(.data[[input$var]] > !!input$min)
```


