# Uploads and Downloads

```{r}
library(tidyverse)
library(shiny)
library(magrittr)
```

# Upload 

## UI

```{r}
ui <- fluidPage(
    fileInput("upload", "Upload a File",
              multiple = TRUE,
              accept = c("text/csv",
                         "text/comma-separated-values,text/plain",
                         ".csv"),
              buttonLabel = "Input",
              placeholder = "Upload a dataset"),
    tableOutput("files"))
```



## Server

Most inputs use simple vectors, but input$file returns a dataframe with 4 columns:

- name : the original file name 
- size : the file size in bytes. By default the user can only upload files up to 5 MB. This can be increased by setting options prior to starting shiny. For example, to allow up to 10MB run options(shiny.maxRequestSize = 10 * 1024^2)
- type : the MIME type of the file. 
- datapath : the path to where the data has been uploaded on the server. The data is always saved to a temporary directory and given a temporary name.

```{r}
server <- function(input, output, session) {
    output$files <- renderTable(input$upload)}

shinyApp(ui, server)
```

## Uploading Data

If the user is uploading a dataset, there are three details to be aware of:

- input$file is initialized to null on page load, so you will need req(input$file) to make sure your code waits until the first file is uploaded
- The accept argument allows you to limit the possible inputs. This is only a suggestion to the browser and is not always enforced, so it is good practice to also validate it yourself

Here is an app that allows the user to upload a csv file and see the first n rows

```{r}
library(DT)

ui <- fluidPage(
    fileInput("file", NULL, accept = c(".csv", ".tsv")),
    numericInput("n", "Rows", value = 5, min = 1, step = 1),
    dataTableOutput("head"))

server <- function(input, output, session) {
    data <- reactive({
        req(input$file)

        ext <- tools::file_ext(input$file$name)

        switch(ext,
               csv = vroom::vroom(input$file$datapath, delim = ","),
               tsv = vroom::vroom(input$file$datapath, delim = "\t"),
               validate("Invalid File; Please upload a .csv or .tsv file"))})

    output$head <- renderDataTable(head(data(), input$n))
}

shinyApp(ui, server)
```

# Download 

## Basics 

You can customize the appearance of the download button by using the class argument with one of : 

- "btn-primary"
- "btn-success"
- "btn-info"
- "btn-warning"
- "btn-danger"

and the size with : 

- "btn-lg"
- "btn-sm"
- "btn-xs"

or space the entire width of the element with : 

- "btn-block"

```{r}
ui <- fluidPage(
    downloadButton("download1", class = "btn-success"),
    downloadLink("download2"))
```

Unlike other outputs, downloadButton is not paired with a render function - instead you use downloadHandler which looks something like this:

```{r}
output$download <- downloadHandler(
    filename = function() {paste0(input$dataset, ".csv")},
    content = function(file) {write.csv(data(), file)}
)
```

downloadHandler has two arguments, both functions:

- filename should be a function with no args that returns a file name as a string. This is used to create the name that will be shown to the user in the download dialog box
- content should be a function with one argument, file, which is the path to save the file. This is used to save the file in a place that shiny knows about so it can be sent to the user.

## Downloading Data

The following app shows off the basics of data download by allowing you to download any dataset in the datasets package as a TSV file. 

```{r}
library(vroom)

ui <- fluidPage(
    selectInput("dataset", "Pick a dataset", ls("package:datasets")),
    dataTableOutput("preview"),
    downloadButton("download", "Download .tsv"))

server <- function(input, output, session) {
    data <- reactive({
        out <- get(input$dataset, "package:datasets")

        if (!is.data.frame(out)) {
            validate(paste0("'", input$dataset, "' is not a data frame"))
        }

        out})

    output$preview <- renderDataTable(head(data()))

    output$download <- downloadHandler(filename = function() paste0(input$dataset, ".tsv"),
                                       content = function(file) vroom_write(data(), file))
}

shinyApp(ui, server)
```

Note the use of validate() to only allow the user to download datasets that are dataframes. A better approach would be to prefilter the list.

## Downloading Reports 

As well as downloading data, we may want users of the app to be able to download a report that summarizes the result of interactive exploration in the shiny app.

One way to achieve this is with a parameterized rmarkdown document: 

bookdown.org/yihui/markdown/parameterized-reports.html

A parameterized rmd file has a params field in the YAML metadata:

```{yaml}
title: My Document 
output: html_document 
params:
    year: 2020
    region: Europe 
    printcode: TRUE
    data: file.csv
```

and inside the document we can refer to these values using params$year, params$region, etc. 

The values in the YAML metadata are defaults -- they can be over-ridden by providing the params argument in a call to rmarkdown::render(). This makes it easy to generate many different reports from the same .Rmd

```{r}
ui <- fluidPage(
    sliderInput("n", "Number of Points", 1, 100, 50),
    downloadButton("report", "Generate Report"))

server <- function(input, output, session) {
    output$report <- downloadHandler(
        filename = "report.html",
        content = function(file) {
            params <- list(n = input$n)
            rmarkdown::render("report.Rmd",
                              output_file = file,
                              params = params,
                              envir = new.env(parent = globalenv()))})}

shinyApp(ui, server)
```

Some tricks worth noting:

- If the report takes time to generate, use some of the techniques from ch 7 to let the user know 
- In many deployment scenarios you won't be able to write to the working directory, which rmarkdown will attempt to do. A workaround is copying the report to a temp directory when the app starts 

```{r}
report_path <- tempfile(fileext = ".Rmd")
file.copy("report.Rmd", report_path, overwrite = TRUE)
```

Then replace "report.Rmd" with report_path in the call to rmarkdown::render()

- By default, rmarkdown will render the report in the current process which means that it will inherit settings from the shiny app like loaded packages and options. For greater robustness, its recommended to run render in a separate R session using callr:

```{r}
render_report <- function(input, outpout, params) {
    rmarkdown::render(input,
                      output_file = output,
                      params = params,
                      envir = new.env(parent = globalenv()))}

server <- function(input, output, session) {
    output$report <- downloadHandler(
        filename = "report.html",
        content = function(file) {
            params <- list(n = input$slider)
            callr::r(render_report,
                     list(input = report_path, output = file, params = params))})}
```


# Case Study 

Upload a file -> preview it -> perform transformations -> let user download

### Uploading and Parsing

```{r}
ui_upload <- sidebarLayout(
    sidebarPanel(
        fileInput("file", "Data", buttonLabel = "Upload"),
        textInput("delim", "Delimiter (leave blank to guess)", ""),
        numericInput("skip", "Rows to Skip", 0, min = 0),
        numericInput("rows", "Rows to Preview", 10, min = 1)),
    mainPanel(
        h3("Raw Data"),
        tableOutput("preview1")))
```

### Cleaning the File

```{r}
ui_clean <- sidebarLayout(
    sidebarPanel(
        checkboxInput("snake", "Rename columns to snake_case?"),
        checkboxInput("constant", "Remove constant columns?"),
        checkboxInput("empty", "Remove empty columns?")),
    mainPanel(
        h3("Cleaner Data"),
        tableOutput("preview2")))
```

### Downloading the File

```{r}
ui_download <- fluidRow(
    column(width = 12, downloadButton("download", class = "btn-block")))
```

### Assemble 

```{r}
ui <- fluidPage(
    ui_upload,
    ui_clean,
    ui_download
)
```

### Server

```{r}
server <- function(input, output, session) {
    # upload
    raw <- reactive({
        req(input$file)
        delim <- if (input$delim == "") NULL else input$delim
        vroom::vroom(input$file$datapath, delim = delim, skip = input$skip)})

    output$preview1 <- renderTable(head(raw(), input$rows))

    # clean
    tidied <- reactive({
        out <- raw()
        if (input$snake) { names(out) <- janitor::make_clean_names(names(out)) }
})
}
```
